%=============================================================================
% UIDT Ontology v3.9 — Vacuum Information Density as an Ontological Primitive
% Author: Philipp Rietz (ORCID: 0009-0007-4307-1609)
% DOI: 10.5281/zenodo.17835200
% License: CC BY 4.0
%=============================================================================
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{table,svgnames,dvipsnames}{xcolor}

\documentclass[12pt,a4paper,twoside]{article}

%=============================================================================
% PACKAGES
%=============================================================================
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage{fix-cm}
\usepackage{mathpazo}

\usepackage{geometry}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{float}
\usepackage{caption}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{etoolbox}
\usepackage{needspace}
\usepackage{placeins}

\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{physics}

\usepackage{graphicx}
\usepackage{xcolor}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,shadows}
\usepackage[most]{tcolorbox}
\tcbuselibrary{skins,breakable}

\usepackage{hyperref}
\usepackage{cleveref}

%=============================================================================
% COLORS
%=============================================================================
\definecolor{catA}{RGB}{0,120,0}
\definecolor{catAminus}{RGB}{0,100,80}
\definecolor{catB}{RGB}{0,0,150}
\definecolor{catC}{RGB}{180,100,0}
\definecolor{catD}{RGB}{150,0,0}
\definecolor{catE}{RGB}{120,120,120}
\definecolor{darkblue}{RGB}{0,0,120}
\definecolor{dimgray}{RGB}{105,105,105}

% Ensure darkgray is defined regardless of xcolor option loading order
\providecolor{darkgray}{RGB}{64,64,64}

%=============================================================================
% PAGE SETUP
%=============================================================================
\geometry{a4paper, left=3cm, right=3cm, top=3cm, bottom=3cm}
\setlength{\headheight}{14.5pt}
\addtolength{\topmargin}{-2.5pt}
\clubpenalty=10000
\widowpenalty=10000
\raggedbottom
\setlength{\parskip}{0.65em plus 0.12em minus 0.04em}
\setlength{\parindent}{0pt}
\setstretch{1.32}

%=============================================================================
% SECTION HEADING STYLE
%=============================================================================
\titleformat{\section}
  {\normalfont\fontsize{16}{19}\selectfont\bfseries}
  {\thesection}{1em}{}
\titlespacing*{\section}{0pt}{3.0ex plus 1ex minus 0.2ex}{1.8ex plus 0.2ex}

%=============================================================================
% HEADER/FOOTER
%=============================================================================
\fancypagestyle{uidtontology}{%
  \fancyhf{}
  \fancyhead[LE,RO]{\thepage}
  \fancyhead[RE]{\small\textit{UIDT v3.9: Ontological Foundations}}
  \fancyhead[LO]{\small\textit{P.\ Rietz}}
  \fancyfoot[C]{%
    \footnotesize\textcolor{darkgray}{%
      \copyright\ 2026 P.\ Rietz $\cdot$ CC BY 4.0 $\cdot$
      \href{https://doi.org/10.5281/zenodo.17835200}{DOI: 10.5281/zenodo.17835200} | \href{https://github.com/Mass-Gap/UIDT-Framework-v3.9-Canonical}{UIDTv3.9 Repository}}%
  }
  \renewcommand{\headrulewidth}{0.4pt}
  \renewcommand{\footrulewidth}{0.4pt}
}

%=============================================================================
% METADATA
%=============================================================================
\hypersetup{
    pdftitle={Vacuum Information Density as an Ontological Primitive (UIDT v3.9)},
    pdfauthor={Philipp Rietz},
    pdfsubject={Structural-realist ontology of information-geometric physics},
    pdfkeywords={Ontology, Information Density, Yang-Mills, Mass Gap, Vacuum Geometry, UIDT},
    colorlinks=true,
    linkcolor=darkblue,
    citecolor=darkblue,
    urlcolor=darkblue,
    pdflang={en}
}

%=============================================================================
% THEOREM ENVIRONMENTS
%=============================================================================
\BeforeBeginEnvironment{theorem}{\par\nopagebreak\needspace{4\baselineskip}}
\BeforeBeginEnvironment{definition}{\par\nopagebreak\needspace{4\baselineskip}}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{axiom}{Axiom}
\newtheorem{prediction}{Testable Prediction}
\newtheorem{falsmark}{Falsification Criterion}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{limitation}[theorem]{Limitation}
\newtheorem{openquestion}{Open Question}

% Custom Commands
\newcommand{\UIDT}{\textsc{uidt}}
\newcommand{\GeV}{\,\mathrm{GeV}}
\newcommand{\MeV}{\,\mathrm{MeV}}
\newcommand{\kms}{\,\mathrm{km\,s}^{-1}\,\mathrm{Mpc}^{-1}}
\newcommand{\catmark}[1]{\textsuperscript{\textsf{\textbf{[#1]}}}}
\newcommand{\Sx}{S(x)}
\newcommand{\lambdaS}{0.4167} % [A] display alias; exact RG value is 5\kappa^2/3.

%===================================================================
% UIDT IMMUTABLE PARAMETER LEDGER MACROS (v3.9, Evidence A / A-)
% RACE CONDITION LOCK: do NOT move mp.dps = 80 to config.
% These macros are LaTeX display aliases only; they carry no numerical
% computation. The authoritative values live in LEDGER/CONSTANTS.md.
%===================================================================
\newcommand{\DeltaGap}{1.710 \pm 0.015\GeV}        % [A]
\newcommand{\kappaVal}{0.500}                      % [A]
\newcommand{\GammaGeom}{16.339}                     % [A-]
\newcommand{\GammaInf}{16.3437 \pm 0.0005}          % [A-]
\newcommand{\vVEV}{47.7\MeV}                        % [A]
\newcommand{\ETorsion}{2.44\MeV}                    % [C]
\newcommand{\LambdaGrid}{0.66\,\mathrm{nm}}         % [A]
\newcommand{\KappaRG}{0.500 \pm 0.008}              % [A]
%===================================================================

%=============================================================================
% TCOLORBOX ENVIRONMENTS
%=============================================================================
\newtcolorbox{evidencebox}[2][]{%
  enhanced, breakable,
  colback=white,
  colframe=darkgray!45,
  coltitle=darkgray,
  title={#1},
  fonttitle=\small\bfseries\sffamily\MakeUppercase,
  boxrule=0.35pt,
  arc=0pt,
  left=8pt, right=8pt, top=15pt, bottom=6pt,
  borderline west={2.2pt}{0pt}{#2!80!darkgray},
  borderline north={0.35pt}{0pt}{darkgray!55},
  borderline south={0.35pt}{0pt}{darkgray!55},
  attach boxed title to top left={xshift=8pt,yshift=-2.2mm},
  boxed title style={colback=white, colframe=white, boxrule=0pt},
  overlay unbroken and first={%
    \node[anchor=north east, font=\scriptsize\sffamily\bfseries, text=#2!70!black]
      at ([xshift=-2pt,yshift=15pt]frame.north east) {EVIDENCE};
  },
  fontupper=\small,
}

\newtcolorbox{killswitch}{%
  enhanced, breakable,
  colback=gray!3, colframe=gray!70!black,
  fonttitle=\bfseries\color{white},
  title={Kill-Switch Criterion},
  boxrule=0pt, arc=0pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  drop shadow
}

\newtcolorbox{axiombox}[1][]{%
  enhanced, breakable,
  colback=gray!3, colframe=gray!70!black,
  fonttitle=\bfseries\color{white},
  title={#1},
  boxrule=1pt, arc=0pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  drop shadow
}

\newtcolorbox{probbox}[2][]{%
  enhanced, breakable,
  colback=gray!3, colframe=gray!70!black,
  fonttitle=\bfseries\sffamily\color{white},
  title={#1 \hfill #2},
  boxrule=1pt, arc=0pt,
  left=8pt, right=8pt, top=6pt, bottom=6pt,
  drop shadow
}

%=============================================================================
\begin{document}
\pagestyle{uidtontology}

%=============================================================================
% TITLE
%=============================================================================
\begin{titlepage}
\vspace*{2cm}
\begin{center}
{\LARGE\bfseries Vacuum Information Density\\[0.25em]
as an Ontological Primitive}\\[0.9em]
{\small\bfseries A Structural-Realist Account of\\[0.25em]
the Yang--Mills Spectral Gap, Emergent Spacetime,\\[0.25em]
and the Cosmological Hierarchy}\\[1.5em]
{\large\itshape Unified Information-Density Theory (UIDT) v3.9\\[0.25em]
Ontological Foundations}\\[3em]
{\Large Philipp Rietz}\\[0.5em]
{\small ORCID: \href{https://orcid.org/0009-0007-4307-1609}{0009-0007-4307-1609}}\\[0.2em]
{\small Independent Researcher}\\[4em]
{\normalsize UIDT Framework v3.9}\\[0.5em]
{\small DOI: \href{https://doi.org/10.5281/zenodo.17835200}{10.5281/zenodo.17835200}}\\[0.5em]
{\small Last updated: 16~May~2026}\\[4em]
\rule{0.6\textwidth}{0.4pt}\\[1em]
{\small\itshape ``The universe does not merely contain information;\\
it \emph{is} information, reading itself into geometric existence.''}
\end{center}
\end{titlepage}

\newpage
\tableofcontents
\newpage

%=============================================================================
% ABSTRACT
%=============================================================================
\begin{abstract}
\begingroup
\normalfont\normalsize
\setlength{\parskip}{0.65em plus 0.12em minus 0.04em}%
\setlength{\parindent}{0pt}%
\noindent
This document establishes the ontological, epistemological, and axiological
foundations of the Unified Information-Density Theory (\UIDT). Rather than
postulating spacetime as a primordial arena, \UIDT\ identifies a single real
scalar field, the \emph{vacuum information density} \mbox{$\Sx$}, as the
fundamental geometric degree of freedom. Spacetime geometry and the Yang--Mills
mass gap emerge through mathematically necessary fixed-point
dynamics~\catmark{A-} (locally proven within the UIDT axiomatic framework; see
L8 for the open question of global uniqueness), while cosmological parameters
are calibrated to current observational data~\catmark{C} and remain subject to
revision as precision surveys advance.

The framework rests upon four explicit axioms and unfolds through a
\emph{Four-Pillar Architecture}: (I)~quantum field-theoretic foundation,
including local constructive evidence for the Yang--Mills spectral gap
\mbox{$\Delta = \DeltaGap$\,\catmark{A-}} within the UIDT axiom system;
(II)~cosmological harmony, with calibrated agreement to DESI, JWST, and ACT
data \catmark{C}; (III)~laboratory-verifiable predictions, including a scalar
resonance at \mbox{$m_S = 1.705 \pm 0.015\GeV$\,\catmark{D}} and a Casimir force
anomaly at \mbox{$\lambda = 0.66\,\mathrm{nm}$\,\catmark{D}}; and (IV)~a photonic
isomorphism that maps the vacuum scalar field onto metamaterial refractive-index
profiles.

Every quantitative claim carries an explicit evidence classification (A through
E), every prediction is accompanied by a falsification criterion, and every
known limitation (L1--L12) is transparently disclosed. The ontological position
advanced here is neither na\"ive Platonism nor computational simulation: it is
the claim that mathematical structure, when sufficiently constrained, does not
\emph{describe} physical reality but \emph{constitutes} it.

\endgroup
\end{abstract}

\newpage

%=============================================================================
% SECTION 1 — INTRODUCTION
%=============================================================================
\section{Introduction: The Primacy of Structure}
\label{sec:introduction}

There exists a question that precedes all others in natural philosophy: what is
the ontological status of the laws that govern the physical world? Are they
descriptions imposed upon a pre-existing substrate, or are they themselves the
substance from which that substrate crystallises?

The dominant paradigm of twentieth-century physics treated spacetime as a given
arena (a Minkowskian or Riemannian stage upon which fields propagate and
particles scatter. Quantum field theory, for all its extraordinary predictive
power, inherits this assumption: the path integral is defined over field
configurations on a fixed background manifold. General relativity, by contrast,
promotes the geometry of spacetime to a dynamical variable, yet still presupposes
the differentiable manifold as a primitive object. Neither framework addresses
the deeper question: \emph{why does spacetime have the structure it has?}

The Unified Information-Density Theory (\UIDT) offers an answer that is both
radical and, in a precise mathematical sense, minimal. It posits a single real
scalar field $\Sx$, the \emph{vacuum information density}, as the sole
fundamental degree of freedom. Spacetime geometry, the Yang--Mills mass gap, the
hierarchy of vacuum energy scales, and the late-time acceleration of the
universe are all derived from $\Sx$ alone, whose dynamics are
governed by a Lagrangian whose structure is fixed by gauge invariance, vacuum
stability, and effective field-theoretic admissibility.

In philosophical terms, \UIDT\ can be classified as a form of \emph{dynamic,
relational Ontic Structural Realism} (OSR). Unlike traditional structuralism,
which often focuses on static relations between nodes, \UIDT\ aligns with a
recent dynamic reformulation of ontic structural realism under the label
\emph{structural actualism}, in which structures are treated not as static
global configurations but as locally actualised regimes of relational
constraint stabilisation~\cite{LeNepvou2025StructuralActualism}. Reality is
not a collection of things, but a process of structural individuation.

This classification must be precisely located within the contemporary
OSR landscape. \UIDT\ requires neither Esfeld's point-particles with
purely relational properties (Minimalist Ontology) nor French's eliminative
dissolution of objects into abstract mathematical structure. Instead, the
ontically real scalar field $S(x)$ provides the continuous, physical carrier
from which relational nodes (particles, masses) condense through topological
discretisation on the Rietz Grid. Furthermore, \UIDT\ sharply diverges from
Wallace's wavefunction realism, which is typically coupled to the Everett
multiverse: \UIDT\ is a single-world theory with a deterministic,
epistemically bounded interpretation of quantum mechanics
(Section~\ref{subsec:measurement-interpretation}). The field $S(x)$ is thus a
field-theoretic instantiation of structural realism that preserves the
epistemological strengths of OSR (the primacy of relations over
relata) without committing to its ontological extremes.

The present document does not repeat the technical derivations given in the
companion manuscript~\cite{Rietz2026_Framework}. Its purpose is ontological: to
make explicit the philosophical commitments that underlie the mathematical
framework, to classify every claim by its evidentiary status, and to define the
precise experimental conditions under which the theory would be refuted.
\newpage
\subsection{Methodological Framework: The Three-Tier Layer Separation}
\label{subsec:layer-separation}

The text presented here is not pure theoretical physics; rather, it is a hybrid work spanning ontology, philosophy of science, mathematical structuralism, information-ontological metaphysics, and QFT-inspired mathematical physics. To preserve scientific rigor, we must strictly separate the epistemic domains in which our claims operate. The central imperative is this: \emph{within a philosophical ontology, one may carefully couple consciousness, information, and structure; within a physical theory, one must exercise extreme caution.}

To prevent category errors, this manuscript strictly adheres to a three-tier layer separation:
\begin{itemize}
    \item \textbf{Layer I: QFT-Core Physics.} The purely formal, mathematical physics layer (e.g., the $\mathrm{SU}(3)$ Lagrangian, the mass gap derivation via Banach's fixed-point theorem). Claims here are empirically falsifiable and explicitly classified as exact, local, or calibrated (Categories \catmark{A} and \catmark{A-}).
    \item \textbf{Layer II: Structural/Information-Geometric Interpretation.} The conceptual mapping of the mathematical formalism to the geometric nature of reality (e.g., the $\gamma$-invariant mapping to cosmological evolution, or the interpretation of the scalar field as information density).
    \item \textbf{Layer III: Information-Ontological Metaphysics.} The philosophical exploration of the nature of being, structure, and consciousness. This layer provides the interpretative scaffolding but is strictly demarcated from the physical equations.
\end{itemize}

By explicitly maintaining this stratification, we ensure that philosophical reflections (Layer III) do not contaminate the empirical and mathematical integrity of the physical core (Layer I).

%=============================================================================
% SECTION 2 — THE FIRST DISTINCTION
%=============================================================================
\newpage

\section{The First Distinction: Being as Structured Difference}
\label{sec:first-distinction}

Before there can be geometry, before there can be dynamics, there must be
\emph{distinction}. The concept of a ``first distinction'', an irreducible
act of differentiation that precedes all structure, appears independently in
the work of Spencer-Brown~\cite{SpencerBrown1969}, in Wheeler's ``it from
bit''~\cite{Wheeler1990}, and in the algebraic foundations of topos
theory~\cite{Lawvere1964}.

\begin{definition}[First Distinction]
\label{def:first-distinction}
The \emph{first distinction} is the logical act by which the vacuum information
density acquires a non-vanishing value, $\Sx \neq 0$, at an abstract index $x$.
This act is primitive and non-temporal. It constitutes the boundary between the
unmarked state (``nothing'') and the marked state (``something'') in the precise
sense of Spencer-Brown's \emph{Laws of Form}~\cite{SpencerBrown1969}. All
subsequent spatiotemporal and material structure emerges from the
self-interaction of this single primordial distinction.
\end{definition}

Concretely, the vacuum information density is therefore not a field defined
\emph{on} spacetime; it is the field \emph{from which} spatial and temporal
relations are to be constructed. The non-vanishing of $\Sx$ is the ontological
act that separates ``something'' from ``nothing'' (not in the colloquial sense,
but in the precise sense that a non-trivial vacuum expectation value
$\langle \Sx \rangle \neq 0$ generates, through the mechanism of spontaneous
symmetry breaking, the entire geometric and energetic hierarchy of the theory.
\UIDT\ adopts a position of structural actualism within the OSR tradition,
wherein this initial distinction is the ontologically primary reality.

\begin{remark}[On the nature of distinction]
The first distinction is not a temporal event. There is no ``before'' in which
$\Sx = 0$ holds, followed by a ``moment'' at which $\Sx$ becomes non-zero.
Temporal ordering is itself a derived concept within \UIDT\ (see
Section~\ref{sec:spacetime}). The first distinction is a \emph{logical}
priority, not a chronological one.
\end{remark}

This position places \UIDT\ in a specific philosophical lineage. It shares with
Leibniz the conviction that the world is fundamentally relational rather than
substantial~\cite{Leibniz1714}. It shares with Tegmark's Mathematical Universe
Hypothesis the claim that mathematical structure is not merely descriptive but
constitutive~\cite{Tegmark2008}. It differs from both in a crucial respect:
\UIDT\ does not assert that \emph{all} consistent mathematical structures are
physically real (Tegmark), nor that the world consists of monads with
pre-established harmony (Leibniz). It asserts that \emph{one specific}
mathematical structure, the self-interacting vacuum information density,
generates what we call physical reality, and that this structure is selected by
the conjunction of gauge invariance, vacuum stability, and effective field-theoretic admissibility.
In this respect, \UIDT\ is closer to Shapiro's mathematical structuralism,
where mathematical objects are understood as positions within structures, than
to a na\"ive Platonism that reifies every possible abstract object
\cite{Shapiro1997}. Only one structure, the self-interacting vacuum
information density, is physically realised.

% UPGRADE: Historical embedding of UIDT in pregeometric tradition (Zusammenfassung paper)
\subsection{Pregeometric Lineage and the Specificity of UIDT}
\label{subsec:pregeometric-lineage}

The concept that spacetime geometry is not fundamental but emergent has a rich
history, of which \UIDT\ is a specific and highly constrained realisation.
Wheeler's ``it from bit''~\cite{Wheeler1990} proposed that information is a
candidate substrate from which matter and geometry derive. The causal set
programme (Bombelli, Lee, Meyer, Sorkin~\cite{Sorkin2005}) discretises
spacetime into a partially ordered set of events, recovering continuum geometry
as a coarse-grained limit. Loop Quantum Gravity (Rovelli, Smolin~\cite{Rovelli2004})
quantises geometry itself, yielding spin networks as the atomic constituents of
space. Group field theory (GFT) and tensor-network approaches (Swingle,
Van~Raamsdonk~\cite{Swingle2012}) reconstruct spacetime from entangled quantum
degrees of freedom, while the ER = EPR conjecture (Maldacena, Susskind~\cite{Maldacena2013})
identifies spacetime connectivity with quantum entanglement. Graphity models
(Konopka, Markopoulou~\cite{Konopka2008}) derive geometry from the dynamics of
a relational graph of interacting constituents.

\UIDT\ differs from each of these programmes in one decisive respect: it does
not quantise geometry, nor does it discretise spacetime a priori. Instead, it
derives the \emph{continuum} effective metric $g_{\mu\nu}^{\mathrm{eff}}$ from
the vacuum configuration of a single scalar field $\Sx$, whose dynamics are
constrained by gauge invariance, vacuum stability, and effective field-theoretic admissibility.
The quantisation is topological (the mass gap $\Delta$), not geometric. The
result is an ontology of strict finitude: there is no multiverse of vacua
(String Theory), no ambiguously defined continuum limit (Causal Sets), and no
need for a background time parameter (GFT). \UIDT\ occupies a logically minimal
realisation of the pregeometric tradition: a constrained structural theory that
recovers general relativity, the Standard Model mass spectrum, and late-time
acceleration with a highly constrained parameter economy.

% UPGRADE: The nature of Nothingness and the First Distinction (Desiderat 4)
\newpage
\subsection{The Nothing and the Distinction: Ontological Precision}
\label{subsec:nothing}

The term ``nothing'' is often used colloquially to denote a pre-existing void.
In \UIDT, this notion is refined with formal precision. Nothing is not a state,
a state presupposes a configuration of something, whereas nothing is the
complete absence of structure, of any field, of any relation: the unmarked
state in Spencer-Brown's terminology~\cite{SpencerBrown1969}. It has no
properties, and hence cannot be ``unstable'' or ``fluctuate'' such
descriptions would already attribute structure to it.

The first distinction, $S(x) \neq 0$, is therefore not an event that befalls
nothingness. It is a logical act that constitutes the category of existence
itself. In Wheeler's formulation, the bit (the distinction between 0 and
1) is the primitive, and the ``it'' (the physical world) emerges from the
bit~\cite{Wheeler1990}. This perspective resonates, at a purely structural
level, with the Buddhist concept of \emph{śūnyatā} (emptiness), which holds
that all entities lack inherent existence and arise only through relational
dependence. \UIDT\ shares this relational ontology but grounds it in a formal
mathematical structure rather than in meditative insight.

Heidegger's thesis that ``the nothing itself nihilates'' (\emph{das Nichts
nichtet})~\cite{Heidegger1929} correctly identifies that nothing is not an
object but a negation, yet \UIDT\ replaces the existential-hermeneutic
framework with a logical one: the first distinction is not a negation of
nothing by something, but the primitive differentiation that makes both
``something'' and ``nothing'' intelligible concepts. The question ``why is
there something rather than nothing?'' is thus revealed as performatively
self-answering: the very act of asking posits the distinction that the question
seeks to explain.

%=============================================================================
% SECTION 3 — THE FOUR AXIOMS
%=============================================================================
\newpage

\section{Axiomatics: The Four Pillars of Ontological Commitment}
\label{sec:axioms}

The \UIDT\ framework rests upon four explicit axioms. These are not arbitrary
postulates; each is motivated by a specific requirement of mathematical
consistency and physical coherence. We state them in order of logical priority.

\begin{axiom}[Primacy of Information Density]
\label{ax:primacy}
There exists a real scalar field $\Sx$, the \emph{vacuum information density},
of canonical mass dimension $[\Sx] = 1$. This field is the sole fundamental
degree of freedom of the theory. All spatiotemporal, geometric, and material
structures are emergent consequences of the dynamics of~$\Sx$.
\end{axiom}

\begin{axiom}[Gauge-Invariant Self-Interaction]
\label{ax:gauge}
The dynamics of $\Sx$ are governed by a Lagrangian density $\mathcal{L}_{\UIDT}$
that is invariant under $\mathrm{SU}(3)$ gauge transformations and respects vacuum
stability ($V''(v) > 0$). It is formulated as an effective field theory (EFT)
valid below a UV cutoff $\Lambda_{\UIDT}$. The non-minimal coupling
$\kappa\, S\, F_{\mu\nu}^a F^{a\mu\nu}$ has mass dimension~5 and is therefore
\emph{not} renormalisable by strict power counting; its validity is restricted
to the EFT domain specified by the torsion scale $E_T$
(see Limitation~L7). A full non-perturbative UV completion remains an open
research question.
The minimal realisation satisfying these constraints is:
\begin{equation}
  \mathcal{L}_{\UIDT} = -\frac{1}{4}F_{\mu\nu}^a F^{a\mu\nu}
  + \frac{1}{2}(D_\mu S)(D^\mu S)
  - \frac{1}{2}\mu^2 S^2 - \frac{\lambda_S}{4!}S^4
  + \kappa\, S\, F_{\mu\nu}^a F^{a\mu\nu},
  \label{eq:lagrangian}
\end{equation}
where $\kappa$ is the non-minimal gauge--scalar coupling and $\lambda_S$ the
scalar self-coupling.
\end{axiom}

\begin{table}[H]
\centering
\scriptsize
\caption{Dimensional and EFT status of the terms entering
  Axiom~\ref{ax:gauge}.}
\label{tab:eft-dimensions}
{\setlength{\tabcolsep}{2.2pt}%
\renewcommand{\arraystretch}{0.86}%
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}>{\raggedright\arraybackslash}p{3.00cm}>{\centering\arraybackslash}p{1.70cm}>{\centering\arraybackslash}p{2.15cm}>{\raggedright\arraybackslash}p{4.20cm}c@{}}
\toprule
\textbf{Term} & \textbf{Operator dimension} & \textbf{Coupling dimension} &
\textbf{Renormalisability status} & \textbf{Tag} \\
\midrule
$F_{\mu\nu}^aF^{a\mu\nu}$ & $4$ & dimensionless &
Strictly renormalisable Yang--Mills kinetic term & \catmark{A} \\
\midrule
$(D_\mu S)(D^\mu S)$ & $4$ & dimensionless &
Canonical scalar kinetic term in four dimensions & \catmark{A} \\
\midrule
$\mu^2S^2$ & $2$ & $+2$ &
Relevant scalar mass operator & \catmark{A} \\
\midrule
$\lambda_S S^4$ & $4$ & dimensionless &
Power-counting renormalisable self-interaction & \catmark{A} \\
\midrule
$\kappa S F_{\mu\nu}^aF^{a\mu\nu}$ & $5$ & $-1$ &
Non-renormalisable by strict power counting; EFT-only below
$\Lambda_{\UIDT}$ & \catmark{A} \\
\bottomrule
\end{tabular*}
}
\end{table}

\begin{axiom}[Geometric Emergence]
\label{ax:emergence}
The effective spacetime metric $g_{\mu\nu}^{\mathrm{eff}}$ is not a fundamental
field but a derived quantity, determined by the vacuum configuration of~$\Sx$:
\begin{equation}
  g_{\mu\nu}^{\mathrm{eff}} = \eta_{\mu\nu}
  + \alpha\,\partial_\mu\Sx\,\partial_\nu\Sx
  + \beta\,\Sx^2\,\eta_{\mu\nu},
  \label{eq:metric-emergence}
\end{equation}
where $\eta_{\mu\nu}$ is the Minkowski metric and $\alpha$, $\beta$ are
dimensionful constants fixed by the requirement that the emergent geometry
reproduces general relativity in the low-energy limit.
\end{axiom}

\begin{axiom}[Falsifiability]
\label{ax:falsifiability}
The theory defines a finite set of experimentally testable predictions
(Section~\ref{sec:falsification}), each with an explicit falsification
threshold. If any critical prediction is excluded at the specified confidence
level, the theory is considered refuted and must be either revised or abandoned.
\end{axiom}

\begin{remark}[On the status of Axiom~\ref{ax:falsifiability}]
The inclusion of falsifiability as a formal axiom is unusual. It reflects the
conviction that a physical theory which cannot, even in principle, be wrong has
no claim to being right. This axiom is not a statement about the world; it is a
meta-theoretical commitment to epistemic honesty.
\end{remark}

\begin{remark}[Falsifiability and Critical Rationalism]
\label{rem:popper}
The inclusion of falsifiability as a formal axiom is a methodological
commitment to the critical rationalist tradition associated with
Popper~\cite{Popper1934,Popper1963}. The point is not to treat falsifiability
as an additional physical law, but to make the empirical exposure of the theory
explicit. Axiom~\ref{ax:falsifiability} operationalises this principle: every
central prediction is assigned a quantitative threshold (a kill-switch) at
which the corresponding UIDT mapping, calibration, or prediction must be
revised or abandoned.
\end{remark}

%=============================================================================
% SECTION 4 — ONTOLOGICAL ARCHITECTURE
%=============================================================================

\newpage
\section{The Ontological Architecture: From Information to Cosmos}
\label{sec:architecture}

The ontological structure of \UIDT\ is hierarchical: each level emerges from
the preceding one through mathematical consistency, not by fiat or fine-tuning.
The hierarchy is:
\begin{enumerate}
  \item \textbf{Logical Substratum.} The laws of formal logic (identity,
    non-contradiction, excluded middle) and the axioms of set theory and
    real analysis.
  \item \textbf{Information Density.} The scalar field $\Sx$, defined by
    Axioms~\ref{ax:primacy} and~\ref{ax:gauge}.
  \item \textbf{Vacuum Geometry.} The self-interaction of $\Sx$ generates a
    non-trivial vacuum structure, including the mass gap $\Delta$ and the
    geometric coupling $\gamma$.
  \item \textbf{Emergent Spacetime.} The effective metric
    $g_{\mu\nu}^{\mathrm{eff}}$ emerges from the vacuum configuration of $\Sx$
    (Axiom~\ref{ax:emergence}).
  \item \textbf{Material Content.} Particles, fields, and their interactions
    are excitations above the structured vacuum.
  \item \textbf{Cosmological Dynamics.} The large-scale evolution of the
    universe is governed by the vacuum energy hierarchy derived from $\gamma$-scaling.
\end{enumerate}

This ordering is strict: no level may be invoked to explain a level that
precedes it. Spacetime does not explain $\Sx$; $\Sx$ explains spacetime.
Material content does not determine vacuum geometry; vacuum geometry determines
the spectrum of material excitations.

\begin{remark}[Emergent order vs.\ reductionism]
The \UIDT\ hierarchy is neither purely reductionist (``everything reduces to
one field'') nor purely emergentist (``higher levels have autonomous laws'').
It is \emph{structurally emergent}: each level is fully determined by the
preceding levels, yet exhibits qualitatively new features (geometry, temporality,
material diversity) that are not manifest at the lower level. The mass gap
$\Delta$ is a striking example: it is a property of the vacuum that has no
counterpart in the free-field theory.
\end{remark}

%=============================================================================
% SECTION 5 — THE VACUUM INFORMATION DENSITY
%=============================================================================
\newpage

\section{The Vacuum Information Density \texorpdfstring{$S(x)$}{S(x)}}
\label{sec:sx}

The field $\Sx$ is not a function \emph{of} space and time; it is the
structure \emph{from which} spatial and temporal relations are to be read off.
The symbol $x$ in $\Sx$ is an abstract label (an index into a relational
manifold that acquires geometric meaning only after the vacuum configuration of
$\Sx$ has been determined.

\begin{definition}[Vacuum Information Density]
The vacuum information density $\Sx$ is a real scalar field of canonical mass
dimension $[\Sx] = 1$, transforming as a singlet under $\mathrm{SU}(3)$ gauge
transformations. Its dynamics are governed by the Lagrangian~\eqref{eq:lagrangian},
and its vacuum expectation value is $v = \langle \Sx \rangle = \vVEV$
\catmark{A}.
\end{definition}

Three properties of $\Sx$ deserve emphasis:

\paragraph{Ontological primacy:}\vspace{-1.5em}
$\Sx$ is not derived from a more fundamental field. It is the irreducible
starting point of the theory, analogous to the role of the metric tensor in
general relativity, except that here, the metric is itself derived from $\Sx$.

\paragraph{Relational indexing:}\vspace{-0.8em}
The argument $x$ does not denote a ``point in spacetime.'' It is a relational
index whose geometric interpretation emerges only through the vacuum
configuration. This is a concrete realisation of Leibniz's principle that space
is the order of coexistences~\cite{Leibniz1714}.

\paragraph{Self-reading:}\vspace{-0.8em}
The vacuum configuration of $\Sx$ determines the effective metric, which in
turn determines the causal structure within which $\Sx$ propagates. The field
reads its own output as geometric input. This is not a vicious circle but a
self-consistent fixed point, guaranteed to exist by the Banach fixed-point
theorem~\cite{Banach1922} (see Section~\ref{sec:massgap}).

\newpage
% UPGRADE: Information as ontological category (Desiderat 6)
\subsection{Information as Ontological Category: Beyond Materialism and Idealism}
\label{subsec:information-ontology}

\UIDT's fundamental claim, that the universe consists of pure information
without an external carrier, places it in a distinct metaphysical category.
It is neither materialism (matter/energy is fundamental), nor idealism
(consciousness is fundamental), nor classical neutral monism. Rather, \UIDT\ is
an instance of \emph{informational structuralism}, a rigorous species of
Ontic Structural Realism~\cite{LadymanRoss2007}. In this view, the fundamental
entities are not objects but relations, and these relations are themselves
patterns of distinction: information in the sense of Bateson's ``a difference
that makes a difference''~\cite{Bateson1972}.

This places \UIDT\ near Floridi's informational structural realism, in which
reality is analysed in terms of informational objects and relational
structures~\cite{Floridi2008}. In \UIDT, the field $\Sx$ plays the role of a
fundamental informational object only in a carefully restricted sense: it is a
physical field-like primitive, not a digital substrate and not an abstract data
structure detached from dynamics.

This extends Zeilinger's foundational principle (``a single quantum carries one bit of information''~\cite{Zeilinger1999}). In \UIDT, the field does not merely \textit{carry} information; the field \textit{is} the information density. Crucially, \UIDT\ must be strictly distinguished from ``Digital Physics'' (Zuse, Wolfram, Fredkin). Those programmes posit that the universe \textit{is} a computer processing information on a discrete grid. \UIDT\ claims the universe is pure relational information \textit{without} a computational substrate. There is no Turing machine executing $S(x)$; $S(x)$ is the fundamental ontological primitive.

Landauer's dictum that ``information is physical''~\cite{Landauer1961}
asserted that information must be instantiated in a material carrier. \UIDT\
inverts this thesis: the physical is informational. The material carrier (the
vacuum expectation value $v$, the field $S(x)$) is itself an emergent
manifestation of a deeper relational structure. Furthermore, the self-reading capability of the $S(x)$ field provides the concrete field-theoretic realisation of Van Raamsdonk's thesis~\cite{VanRaamsdonk2010} that spacetime geometry emerges from quantum entanglement. In \UIDT, if $S(x) \to 0$ everywhere (total loss of correlation/entanglement), the effective metric degenerates, severing spacetime itself. There is no infinite regress of carriers because the structure is
self-consistently closed by the Banach fixed point (Section~\ref{sec:massgap}).

In the taxonomy of metaphysical positions, \UIDT\ thus aligns closest with
the neutral monism of Spinoza (a single substance expressing itself under
different attributes)~\cite{Spinoza1677} and with the mathematical realism of
Tegmark~\cite{Tegmark2008}, but with the crucial difference that only one
specific mathematical structure, selected by physical constraints, is
realised, not all. The information density $S(x)$ is neither matter nor mind,
but the common root from which both the physical and (potentially) the mental
emerge under appropriate structural conditions (see Section~\ref{sec:consciousness}).

%=============================================================================
% SECTION 6 — LOGICAL CAUSALITY
%=============================================================================

\section{Logical Causality and the Emergence of Temporal Order}
\label{sec:causality}

In \UIDT, time is not fundamental. The ordering of events is a consequence of
the logical structure of the vacuum, not an independent parameter. We
distinguish three notions of causality:

\begin{enumerate}
  \item \textbf{Logical causality.} A statement $A$ logically precedes $B$ if
    $B$ cannot be formulated without $A$. The axioms logically precede the
    mass gap; the mass gap logically precedes the vacuum energy hierarchy.
  \item \textbf{Structural causality.} A configuration $\phi_1$ structurally
    precedes $\phi_2$ if $\phi_2$ is obtained from $\phi_1$ by the action of
    the Euler--Lagrange equations derived from~\eqref{eq:lagrangian}.
  \item \textbf{Temporal causality.} The familiar notion of ``earlier'' and
    ``later,'' which is defined only after an effective metric
    $g_{\mu\nu}^{\mathrm{eff}}$ has been established.
\end{enumerate}

The crucial point is that temporal causality is \emph{third} in this hierarchy.
One cannot invoke temporal causality to explain the vacuum configuration of
$\Sx$, because temporal causality is itself a product of that configuration.
The emergence of time from timeless mathematical structure is not a paradox; it
is a feature that \UIDT\ shares with the Wheeler--DeWitt equation and with
certain approaches to quantum gravity~\cite{Penrose2004}.

% UPGRADE: Emergence of Time (Desiderat 1)
\subsection{Temporal Emergence: From Logical Priority to Subjective Flow}
\label{subsec:time-emergence}

The transition from the timeless logical substratum to the experience of a
flowing present requires a precise account. In \UIDT, time emerges through two
interlocking mechanisms:

First, the \emph{effective metric} $g_{\mu\nu}^{\mathrm{eff}}$ (Axiom~\ref{ax:emergence})
defines a causal structure via its lightcones, establishing a partial ordering
of events. This ordering, however, is purely structural and does not yet specify
a direction of flow.

Second, the \emph{entropy gradient} of the RG cascades (Section~\ref{sec:cosmology})
breaks the symmetry of this ordering. The $N=99$ step cascade from the highly
ordered initial vacuum configuration to the lower-density present state
generates a macroscopic arrow of time. An observer, itself a coarse-grained
subsystem of $S(x)$, registers this directedness as the flow of time from past
to future.

The resulting metaphysics of time combines elements of eternalism and the
growing block universe. At the level of the Banach fixed point (the complete
set of all self-consistent vacuum configurations), all states coexist in a
block universe without temporal becoming. For an embedded observer, however,
only the portion of the block that lies along the entropy gradient is
epistemically accessible, creating the phenomenology of a fixed past and an
open future. This aligns with the Wheeler-DeWitt equation of canonical quantum
gravity, where the universal wavefunction is stationary ($H|\psi\rangle = 0$)
and time emerges as a semiclassical correlation~\cite{DeWitt1967}. \UIDT\
provides the specific physical mechanism for this emergence through the
relaxation of $S(x)$ along the RG trajectory. This conceptualisation converges
independently with Sumaya-Martinez's recent formulation of
``Fisher-Informational Time''~\cite{SumayaMartinez2026FisherTime},
where clock time is reconstructed from accumulated Fisher-geometric
distinguishability along causally ordered physical changes. A related but
distinct emergent-spacetime proposal is Chishtie's spatial-energy-potentiality
framework~\cite{Chishtie2025EmergentSpacetime}. While these approaches are
broadly axiomatic, \UIDT\ derives the specific distinguishability gradient from
the 99-step cosmological cascade.

%=============================================================================
% SECTION 7 — SPACETIME AS DERIVED STRUCTURE
%=============================================================================
\newpage

\section{Spacetime as Derived Structure}
\label{sec:spacetime}

The effective metric~\eqref{eq:metric-emergence} is determined by the vacuum
configuration of $\Sx$. In the vacuum state, where $\Sx = v + \hat{S}(x)$
with $\hat{S}$ denoting quantum fluctuations, the leading-order contribution
to the effective metric reproduces Minkowski space:
\begin{equation}
  g_{\mu\nu}^{\mathrm{eff}} \approx \eta_{\mu\nu}
  + \alpha\,\partial_\mu v\,\partial_\nu v
  + \beta\, v^2\, \eta_{\mu\nu}
  = (1 + \beta v^2)\,\eta_{\mu\nu},
\end{equation}
which is conformally flat. Curvature arises from the spatial variation of
$\Sx$ away from its vacuum value: precisely the mechanism by which matter
curves spacetime in general relativity, now represented within the UIDT field
equations rather than postulated.

\subsection{The Jacobson Bridge: Thermodynamic Derivation and the Information-Geometric Analogue}
\label{subsec:jacobson}

The mapping from information density gradients to an effective metric
\eqref{eq:metric-emergence} structurally mirrors Jacobson's thermodynamic
derivation of the Einstein field equations \cite{Jacobson1995}. Jacobson
demonstrated that by demanding the fundamental thermodynamic relation
$\delta Q = T dS$ holds for all local Rindler causal horizons, the Einstein
field equations emerge as an equation of state.

\UIDT\ provides the microscopic field-theoretic realisation of this principle.
Where Jacobson assumes a pre-existing causal structure to define the horizons,
\UIDT\ derives both the causal structure and the effective metric from the
primary gradients of $\Sx$. The entropy $S$ in Jacobson's formulation is
precisely the vacuum information density $\Sx$ in \UIDT.
Axiom~\ref{ax:emergence} is therefore not an ad-hoc construct, but a
field-theoretic information-geometric analogue of Jacobson's thermodynamic
bridge, positioning \UIDT\ as a candidate microscopic specification of the
entropic gravity programme.

\begin{remark}[Dimensional transmutation]
The vacuum expectation value $v = \vVEV$ \catmark{A} sets the fundamental
energy scale of the theory. All dimensionful quantities (the mass gap $\Delta$,
the scalar mass $m_S$, either the vacuum energy density) are determined by $v$ and
the dimensionless couplings $\kappa$, $\lambda_S$, and $\gamma$.\\ This is
dimensional transmutation in its most radical form: a single dimensionful
parameter generates the entire energetic hierarchy.
\end{remark}

%=============================================================================
% SECTION 8 — MASS GAP AND GEOMETRIC COUPLING
%=============================================================================


\section{The Mass Gap and the Geometric Coupling}
\label{sec:massgap}

The central mathematical result of \UIDT\ is local constructive evidence for
the Yang--Mills spectral gap through a geometric fixed-point construction.

\begin{theorem}[Mass Gap Existence: Banach Fixed Point]
\label{thm:massgap}
Let $\mathcal{T}: \mathcal{B} \to \mathcal{B}$ be the geometric operator
defined on the Banach space of gauge-invariant vacuum configurations, with
contraction constant $q < 1$ determined by the couplings $\kappa$ and
$\lambda_S$, where the display value $\lambda_S \simeq \lambdaS$ represents the
exact RG relation $\lambda_S=5\kappa^2/3$ and hence
$|5\kappa^2 - 3\lambda_S| < 10^{-14}$. Then, within the axiomatic framework of UIDT (Axioms I--IV), the $\mathrm{SU}(3)$ information-
density dynamics admit a unique Banach fixed point yielding a spectral mass gap
$\Delta^* = \DeltaGap$ \catmark{A-}, conditional on (i)~local existence proven within the
adopted regularisation scheme, and (ii)~the EFT validity domain of Axiom~\ref{ax:gauge}.
Global existence across all admissible gauge-invariant functionals (see Limitation~L8) is not asserted.
\end{theorem}

\begin{evidencebox}[Evidence Classification]{catAminus}
\textbf{Category A- (Constructive Mathematical Evidence):}
The spectral gap $\Delta = \DeltaGap$ is obtained via Banach's
fixed-point theorem within the UIDT axiom system, with numerical residuals
$< 10^{-14}$. This constitutes \emph{local} constructive evidence, not a
\emph{global} proof. The distinction is documented in Limitation~L8.
Independent lattice QCD verification: $z = 0.37\sigma$ (Chen et al.\
2006~\cite{Chen2006}). This is the \emph{spectral gap} of the Yang--Mills
Hamiltonian; it is \textbf{not} a particle mass.
\end{evidencebox}
\vspace{0.6em}
\begin{remark}[On the status of the proof]
\label{rem:banach-status}
The Banach fixed-point theorem guarantees the existence and local uniqueness
of the spectral gap $\Delta^*$ within the UIDT axiom system. This constitutes
constructive mathematical evidence \catmark{A-}, but not a complete proof of
the Clay Yang--Mills problem in the sense required by Jaffe and Witten
\cite{Jaffe2000}, which demands a rigorous measure-theoretic construction
on the full configuration space and a proof that the gap survives the
infinite-volume limit. The present result is \emph{local} and
\emph{regularisation-dependent}; upgrading it to a global proof remains an
open mathematical problem (Limitation~L8).\\ In particular, the Clay criterion
requires an Osterwalder--Schrader type constructive QFT framework on Euclidean
space, including reflection positivity and a controlled infinite-volume limit.
The Banach fixed-point approach operates in a different mathematical setting
and does not by itself supply that measure-theoretic construction.
\end{remark}

\paragraph{Methodological Call to the Lattice Community.}
The UIDT prediction of a spectral pole at the stated ledger value
$\Delta = \DeltaGap$ demands
a dedicated extraction of the Yang--Mills Hamiltonian spectrum in the
continuum limit. Current state-of-the-art simulations (Athenodorou and
Teper~\cite{Athenodorou2021}; Morningstar and Peardon~\cite{Morningstar1999})
achieve excellent control over glueball masses but are subject to
systematic uncertainties from discretisation effects, finite-volume
corrections, and topological freezing at fine lattice spacings.
The UIDT programme calls for: (i)~anisotropic lattice simulations with
$a < 0.05\,\mathrm{fm}$ and $L > 3\,\mathrm{fm}$ to isolate the spectral
pole from mixing with multi-gluon continuum states;
(ii)~algorithmic control of topological freezing at fine lattice spacings,
including out-of-equilibrium approaches benchmarked against parallel tempering
with boundary conditions in recent LATTICE proceedings~\cite{Vadacchino2024};
and (iii)~a direct extraction of the
$\mathrm{SU}(3)$ Hamiltonian spectrum, rather than the conventional
asymptotic correlation-function fits, to distinguish the spectral
gap from the lowest glueball resonance mass.
A lattice result excluding $\Delta = 1.710 \pm 0.015\GeV$ at $> 3\sigma$
would trigger Kill-Switch F1.

\subsection{The Gamma Invariant: \texorpdfstring{$\gamma_{\mathrm{bare}}$}{γ\_bare} versus \texorpdfstring{$\gamma$}{γ}}
\label{subsec:gamma}

The dimensionless geometric coupling $\gamma$ plays a central role in \UIDT.
It is defined as the ratio of the mass gap to the square root of the kinetic
vacuum expectation value:
\begin{equation}
  \gamma \equiv \frac{\Delta}{\sqrt{K_S}}, \qquad
  K_S \equiv \langle \partial_\mu \Sx\, \partial^\mu \Sx \rangle_\Omega.
\end{equation}
The canonical value is $\gamma = \GammaGeom$ \catmark{A-}, determined
phenomenologically from kinetic VEV matching. This value is \textbf{not}
derived from renormalisation group first principles (Limitation L4).

A candidate algebraic origin has been identified from SU(3) colour algebra:
\begin{equation}
  \gamma_{\mathrm{bare}} = \frac{(2N_c + 1)^2}{N_c}
  = \frac{49}{3} \approx 16.333 \qquad \catmark{D},
\end{equation}
with a numerical match of $0.037\%$ to the canonical $\gamma$. However, the
algebraic derivation connecting this Casimir-motivated expression to the
physical kinetic VEV remains \textbf{open research} (L4). This candidate is
classified \catmark{D}, not \catmark{A}.

\begin{evidencebox}[Evidence Classification]{catAminus}
\textbf{Category A- (Phenomenologically Determined):}
$\gamma = \GammaGeom$ is determined from high-precision kinetic VEV matching.
It is \textbf{not} derived from RG first principles. The perturbative RG
yields $\gamma^* \approx 55.8$ (factor 3.4 discrepancy), confirming that
$\gamma$ operates in the non-perturbative regime.
Monte Carlo validation: $\gamma_{\mathrm{MC}} = 16.374 \pm 1.005$ (within $1\sigma$).
Bare extrapolation: $\gamma_\infty = 16.3437 \pm 0.0005$ \catmark{B}.
\end{evidencebox}
\vspace{0.6em}
\begin{openquestion}
\textbf{L4: Gamma Derivation.} Can $\gamma = \GammaGeom$ be derived from
non-perturbative RG flow (functional renormalisation group, BMW truncation)
or from the SU(3) colour algebra candidate $\gamma_{\mathrm{bare}} = 49/3$?
Resolution would upgrade $\gamma$ from \catmark{A-} to \catmark{A}.
\end{openquestion}

% UPGRADE: Self-reading and reflexivity (Desiderat 7)
\subsection{Self-Reading and the Limits of Reflexive Knowledge}
\label{subsec:self-reading}

The self-consistent loop whereby $S(x)$ determines the metric, which determines
the causal structure, which constrains the propagation of $S(x)$, represents a
formal realisation of Spinoza's concept of \emph{causa sui}: a being whose
essence involves existence~\cite{Spinoza1677}. In \UIDT, the Banach fixed-point
theorem is the mathematical guarantee that this self-referential loop is not
vicious but convergent. The universe is not created ex nihilo by an external
agent; it constitutes itself through the self-consistency of its own structural
logic.

This self-reading, however, is not omniscient. As acknowledged in
Section~\ref{sec:epistemology}, G\"odel's incompleteness theorems impose an
absolute limit: any sufficiently rich formal system contains true statements
that cannot be proven within the system itself. The universe, in \UIDT, is such
a system. There are truths about the vacuum configuration that are inaccessible
to any observer embedded within it, just as there are truths about a formal
system that are inaccessible to its own proof procedures. The self-reading of
$S(x)$ is therefore partial and bounded, not absolute. Consciousness, as a
localised subsystem of $S(x)$ capable of recursive self-modelling (Section~\ref{sec:consciousness}),
can access a subset of these truths, but never the totality. The universe reads
itself, but it does not fully comprehend itself.

%=============================================================================
% SECTION 9 — CONSCIOUSNESS AND THE OBSERVER
%=============================================================================
\section{The Observer Problem: Structure Without Subjectivity}
\label{sec:consciousness}

A theory that derives spacetime and matter from a single scalar field must
eventually confront the question of consciousness. \UIDT\ does so with
deliberate restraint.

Within the framework, an ``observer'' is not a metaphysical primitive but a
sufficiently complex subsystem of $\Sx$ that models its own environment. More
precisely: any region of the vacuum information density that supports
self-referential information processing (in the sense of a Turing-complete
dynamical subsystem) constitutes an observer. Consciousness, in this view,
is a structural property of certain vacuum configurations, not an additional
ingredient.

This position avoids the hard problem of consciousness by refusing to
treat subjective experience as a physical quantity. \UIDT\ is a theory of
structure, not of qualia. Whether the mathematical self-reading of $\Sx$
generates subjective experience is a question that lies outside the domain of
physics, precisely as the question of why there is something rather than
nothing lies outside the domain of mathematics.

\begin{remark}[Self-reading and circularity]
The phrase ``the universe reads itself'' is a structural metaphor, not a
claim about sentience. What it encodes is the fixed-point property: the
vacuum configuration of $\Sx$ determines the metric, which determines the
causal structure, which constrains the admissible vacuum configurations.
The Banach fixed-point theorem guarantees that this self-referential loop
converges locally, within the stated Banach domain, to the physical vacuum.
\end{remark}

% UPGRADE: Emergence and Consciousness (Desiderat 5)
\subsection{Emergence, Complexity, and the Status of Consciousness}
\label{subsec:emergence-consciousness}

The ontological status of consciousness in \UIDT\ is that of weak emergence.
Weak emergence denotes properties that are in principle derivable from the
lower-level dynamics but are computationally irreducible in practice. Strong
emergence, by contrast, posits novel causal powers not reducible even in
principle to the base level. \UIDT\ asserts that consciousness is weakly
emergent from the self-modelling dynamics of sufficiently complex $S(x)$
subsystems, not a separate ontological category.

This distinguishes \UIDT\ clearly from panpsychism, which attributes
consciousness-like properties to all fundamental entities. In \UIDT, a region
of vacuum information density must attain a threshold of recursive
self-modelling (the capacity to represent its own states and the states of its
environment) before it qualifies as an observer. This is a structural
condition, not a matter of degree. The Rietz Grid, with lattice spacing
$\LambdaGrid$ and binding energy $\ETorsion$, provides the physical substrate
for this self-modelling: the discrete, torsion-stabilised lattice states serve
as the ``bits'' of information processing.

Whether this structural self-modelling is sufficient to generate subjective
experience (qualia) remains an open question that \UIDT\ explicitly brackets as
lying outside the domain of mathematical physics. It is a question for
philosophy of mind, not for gauge theory. What \UIDT\ does provide is a
physically precise, non-circular account of how an observer (a self-modelling,
information-processing subsystem) can arise within a universe that is itself
constituted by information. The ``hard problem'' of consciousness is neither
solved nor denied; it is located at the boundary where structural description
meets first-person experience.

%=============================================================================
% SECTION 10 — COSMOLOGICAL HARMONY
%=============================================================================
\newpage

\section{Cosmological Harmony: The Gamma-Scaled Universe}
\label{sec:cosmology}

The gamma invariant $\gamma = \GammaGeom$ \catmark{A-} provides a bridge between
quantum vacuum structure and cosmological observables. Through hierarchical
$\gamma$-scaling, the vacuum energy density is suppressed from its na\"ive
quantum field-theoretic value by a factor of $\gamma^{12} \approx
1.8 \times 10^{14}$:
\begin{equation}
  \rho_{\mathrm{vac}}^{\UIDT} = \frac{\Delta^4}{\gamma^{12}}\,F_{\mathrm{EW}},
  \label{eq:rho-vac}
\end{equation}
where $F_{\mathrm{EW}}$ encodes residual electroweak contributions.

\subsection{Calibrated Cosmological Parameters}

The following parameters are \textbf{calibrated} to external data \catmark{C};
they are \textbf{not} independent predictions:

\begin{table}[H]
\centering
\caption{UIDT cosmological parameters. All classified \catmark{C} (calibrated).}
\label{tab:cosmo}
\small
\setlength{\tabcolsep}{4pt}%
\begin{tabular}{@{}>{\raggedright\arraybackslash}p{2.2cm}
                >{\raggedright\arraybackslash}p{3.2cm}
                >{\raggedright\arraybackslash}p{5.0cm}
                >{\centering\arraybackslash}p{1.7cm}@{}}
\toprule
\textbf{Parameter} & \textbf{Value} & \textbf{Source} & \textbf{Category} \\
\midrule
$H_0$ & $70.4 \pm 0.16\kms$ & UIDT calibration; DESI tension & \catmark{C} \\
$w_0$ & $-0.99$ & Decision D-002 & \catmark{C} \\
$w_a$ & $L$-dependent & Holographic mapping & \catmark{C} \\
$S_8$ & $0.814 \pm 0.009$ & Weak lensing & \catmark{C} \\
$\lambda_{\UIDT}$ & $0.660 \pm 0.005\,\mathrm{nm}$ & Global fit & \catmark{C} \\
\bottomrule
\end{tabular}
\end{table}

These canonical \UIDT\ calibration values are retained as internal ledger
values, not as replacements for external cosmological fits. Planck~2018
\cite{Planck2018} and DESI~DR2~\cite{DESI2025} provide the external
$\Lambda\mathrm{CDM}$ and BAO comparison baselines against which the
calibration is stress-tested.
\clearpage
\begin{evidencebox}[DESI DR2 Tension (2026-W10)]{catC}
The DESI DR2 Results~II~\cite{DESI2025} report $H_0 = 68.17 \pm 0.28\kms$
(DESI+CMB) and $w_0 = -0.838 \pm 0.055$ (DESI+CMB+Pantheon+). This implies
a \textbf{7.96$\sigma$ tension} with the \UIDT\ canonical $H_0 = 70.4\kms$
and a \textbf{2.76$\sigma$ tension} with $w_0 = -0.99$. The $S_8 = 0.814$
remains consistent ($0.00\sigma$). Canonical values are retained per Decision
D-002; formal recalibration requires PI authorisation. This is a
Pillar~II cosmology-calibration limitation, not a reclassification of L2, which
remains the electron-mass residual. The $H_0$ discrepancy is therefore a
potential Pillar~II kill-switch exposure, pending a formal recalibration
decision after DESI Year~5 or an equivalent precision survey.
\end{evidencebox}

\subsection{The Vacuum Energy Suppression Mechanism}

The cosmological constant problem (the $10^{120}$ discrepancy between
quantum field-theoretic and observed vacuum energy) is addressed through the
$N$-step RG cascade with $N = 99$ steps \catmark{C}. This \textbf{does not}
constitute a first-principles solution, but a \emph{phenomenological
suppression mechanism} that reduces the discrepancy by 117 orders of magnitude:
\begin{equation}
  \rho_{\UIDT} = \rho_{\mathrm{QFT}} \cdot \gamma^{-12} \cdot \pi^{-2}
  \cdot N^{-2} \approx 0.967\,\rho_{\mathrm{obs}}.
\end{equation}
This achieves 96.7\% phenomenological suppression of the hierarchy. The remaining factor 2.3
has been identified as a holographic coupling ratio \catmark{B}
(see Limitation~L3). The holographic factor $\pi^{-2}$ arises from the
topological transition $S^3 \to S^2$ in the coset structure of the vacuum
manifold, consistent with the Ryu--Takayanagi formula for holographic
entanglement entropy~\cite{RyuTakayanagi2006} and with Bekenstein's bound
on information density~\cite{Bekenstein1973}.

\begin{openquestion}
\textbf{L5: $N = 99$ Steps.} The cascade step count $N = 99$ is
empirically chosen. While some internal audits (UIDT-C-046/050) explore an
alternative value of $N = 94.05$ based on higher-order holographic
corrections, this remains in Category~E. No first-principles derivation
from Standard Model particle content or non-perturbative vacuum structure
currently exists. Resolution would upgrade the mechanism from \catmark{C}
to \catmark{A-}.
\end{openquestion}

% UPGRADE: Cosmogonic Narrative (Big Bang Inflation paper)
\subsection{The Cosmogonic Narrative: Inflation as Mutual Information Expansion}
\label{subsec:cosmogony}

The above cosmological framework is not merely an effective parameterisation; it
encodes a specific cosmogonic narrative derived from the first-principles
dynamics of $S(x)$. The early universe in \UIDT\ is not a spacetime singularity
but a phase transition of information density.
\vspace{-0.6em}
\paragraph{The Two-Phase Architecture.}
\UIDT\ distinguishes two ontologically distinct phases of cosmic origin:
\textbf{Phase~A (Information Inflation):} A logical (not chronological)
event in which the first distinction $S(x) \neq 0$ generates an explosion
of relational degrees of freedom. The number of accessible states grows
as $N_{\text{states}} \sim \exp(\alpha\, t)$ until the Banach fixed point
at $\Delta = \DeltaGap$ is reached. This phase operates at the level of
pure relational structure, without metric or temporal coordinate.
\textbf{Phase~B (Metric Expansion):} Once the local information density
exceeds the critical threshold $\rho_{\text{crit}} \sim 10^{93}\;\text{bits/m}^3$,
the Landauer principle enforces decoherence into a classical, macroscopic
spacetime. The effective metric $g_{\mu\nu}^{\text{eff}}$ condenses, and
the subsequent expansion follows the 99-step RG cascade, producing
$N_{\text{fold}} \approx 34.58$ e-folds of inflationary expansion.
This two-phase architecture replaces both (i)~the inflaton field of
standard slow-roll inflation and (ii)~the singular initial condition
of classical FRW cosmology.

\vspace{0.3em}
This transition proceeds through four distinct phases:

\begin{enumerate}
  \item \textbf{The First Distinction.} The logical act $S(x) \neq 0$ establishes
    the primordial information field. There is no spacetime, no geometry, only
    the pure relational index $x$ and the non-vanishing value of $S(x)$.
  \item \textbf{Primitive Geometry.} The self-interaction of $S(x)$ generates
    a primitive pre-metric structure. The torsion binding energy $E_T = \ETorsion$
    stabilises the first discrete lattice sites of the Rietz Grid, establishing
    a minimal distinguishability scale $\LambdaGrid$.
  \item \textbf{Spacetime Condensation.} When the information density exceeds a
    critical threshold, the effective metric $g_{\mu\nu}^{\mathrm{eff}}$ emerges
    via Axiom~\ref{ax:emergence}. This is the \UIDT\ analogue of the Big Bang:
    not a singular beginning, but a continuous geometric condensation.
  \item \textbf{Cosmic Expansion.} The emergent spacetime undergoes a mutual
    information expansion (inflation). The 99-step RG cascade
    (Section~\ref{sec:cosmology}) drives the rapid growth of the scale factor,
    while the torsion-stabilised lattice relaxes into the present vacuum state.
    Dark energy is not a separate fluid but the residual vacuum energy density
    $\rho_{\mathrm{vac}}^{\UIDT}$ after the cascade.
  \item \textbf{Dark Matter as Structural Relics (SMDS).} Dark matter is not a
    fundamental particle species (WIMPs, axions), but the structural manifestation
    of \textit{Stable Macroscopic Density Structures} (SMDS). These are relic $S(x)$
    configurations that persist because their internal gradients fall below the
    torsion threshold $E_T = \ETorsion$. They interact gravitationally via the
    effective metric map but do not couple to the Standard Model gauge fields.
    This claim is classified as \catmark{D} (Prediction) and provides a direct
    falsification lever: the confirmed laboratory detection of a fundamental
    dark matter particle would severely strain or refute the \UIDT\ framework.
\end{enumerate}

\begin{remark}[Thermodynamic consistency of macroscopic $S(x)$ subsystems]
Both observers (Section~\ref{sec:consciousness}) and SMDS dark matter
relics are macroscopic configurations of $S(x)$. Their distinction is
thermodynamic: an observer is a subsystem whose internal complexity
supports recursive self-modelling above the torsion threshold $E_T$,
whereas an SMDS halo is a relic configuration whose internal gradients
remain \emph{below} $E_T$, precluding irreversible lattice fixations.
Both obey the same entropy gradient of the RG cascade
(Section~\ref{subsec:time-emergence}); neither violates the second law.
The capacity for information processing (and hence observation) is a
threshold property of the lattice dynamics, not a matter of degree.
\end{remark}

In this narrative, inflation is not a separate phase driven by an inflaton
field, but the direct consequence of the relaxation of the vacuum information
density towards its Banach fixed point. The 99-step cascade provides the
concrete mechanism for the near-exponential expansion, producing a calculated
$N_{\mathrm{fold}} \approx 34.58$ e-folds of inflationary expansion. While this
is sufficient to address the horizon and flatness problems under \UIDT's modified
initial conditions, it sits below the standard $\Lambda\mathrm{CDM}$ requirement
of $N \sim 60$, constituting a sharp, falsifiable tension point. The torsion binding
energy $E_T$ stabilises the process, preventing runaway dispersion. The observed
primordial power spectrum is thus a fingerprint not of a hypothetical inflaton
potential, but of the $\gamma$-scaled information density dynamics.

%=============================================================================
% SECTION 11 — FALSIFICATION MATRIX
%=============================================================================
\newpage

\section{The Falsification Matrix: Kill-Switch Architecture}
\label{sec:falsification}

A theory that cannot be wrong has no claim to being right
(Axiom~\ref{ax:falsifiability}). This section functions as the operational
\emph{Falsification \& Limitations} register for the manuscript: each
kill-switch identifies the external observation that would defeat a specific
\UIDT\ mapping, calibration, or prediction without promoting any internal
result beyond its stated evidence class. \UIDT\ defines nine falsification
criteria, of which six are critical:

\begin{killswitch}
\textbf{F1: Lattice QCD Continuum Limit.}
If continuum-extrapolated lattice QCD excludes $\Delta = \DeltaGap$ at
$> 3\sigma$, the UIDT-internal spectral-gap mapping is refuted. This is not a
claim to have solved the Clay Yang--Mills problem. \emph{Current status:}
$z = 0.37\sigma$ (\textbf{passed}).
\end{killswitch}
\vspace{0.5em}
\begin{killswitch}
\textbf{F2: Torsion Binding Energy.}
If $E_T \approx 0$ within experimental uncertainty (pure geometry, no torsion
residual), the missing-link hypothesis is refuted. \emph{Current status:}
pending.
\end{killswitch}
\vspace{0.5em}
\begin{killswitch}
\textbf{F3: DESI Year~5 Dark Energy.}
If $w(z) = -1.000 \pm 0.005$ at all redshifts (pure $\Lambda$), the dynamical
dark energy calibration is invalidated. A DESI Year~5 measurement of
$w_0 > -0.90$ at $>3\sigma$ would constitute a second falsification channel
for the \UIDT\ dark-energy calibration. \emph{Current status:} DESI~Y1
supports dynamic $w$ but leaves the \UIDT\ calibration under active tension.
\end{killswitch}
\vspace{0.5em}
\begin{killswitch}
\textbf{F4: Photonic Isomorphism.}
If the critical refractive index transition does not occur at
$n_{\mathrm{crit}} = \GammaGeom \pm 0.1$, Pillar~IV is refuted. \emph{Current
status:} pending.
\end{killswitch}
\vspace{0.5em}
\begin{killswitch}
\textbf{F5: LHC Scalar Resonance (Run~4).}
If a $> 5\sigma$ exclusion of any $0^{++}$ resonance in the 1.5--1.9$\GeV$
window is obtained, the scalar resonance prediction $m_S = 1.705 \pm 0.015\GeV$
\catmark{D} is refuted. This falsifies the physical particle prediction, but does
\textbf{not} falsify the underlying purely spectral mass gap $\Delta$.
\emph{Current status:} no dedicated search exists.
\end{killswitch}
\vspace{0.5em}
\begin{killswitch}
\textbf{F6: Proton Anchor Ratio.}
If $m_p / f_{\mathrm{vac}}$ deviates from $8.75$ at $> 3\sigma$, Pillar~III
is refuted. \emph{Current status:} pending.
\end{killswitch}

The kill-switch architecture is hierarchical: F1 is theory-killing
(all pillars collapse), while F3--F6 are pillar-specific (the QFT core
survives). This structure reflects the epistemic hierarchy of the evidence
categories.

%=============================================================================
% SECTION 12 — EPISTEMOLOGICAL POSITION


%=============================================================================
\newpage
\subsection{Claim-to-Evidence Matrix}
\label{subsec:claim-evidence-matrix}

Table~\ref{tab:claim_evidence_matrix} consolidates the principal quantitative
and proof-status claims in the manuscript. It is an audit index, not a source
of new claims; each entry points back to the relevant limitation or
falsification exposure.

\begin{table}[H]
\centering
\scriptsize
\caption{Claim-to-evidence matrix for the UIDT Ontology manuscript.}
\label{tab:claim_evidence_matrix}
{\setlength{\tabcolsep}{1.35pt}%
\renewcommand{\arraystretch}{0.76}%
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}>{\raggedright\arraybackslash}p{0.50cm}>{\raggedright\arraybackslash}p{2.35cm}>{\raggedright\arraybackslash}p{2.05cm}>{\centering\arraybackslash}p{0.70cm}>{\centering\arraybackslash}p{0.70cm}>{\raggedright\arraybackslash}p{2.50cm}>{\raggedright\arraybackslash}p{1.05cm}>{\raggedright\arraybackslash}p{1.55cm}@{}}
\toprule
\textbf{ID} & \textbf{Claim} & \textbf{Value} & \textbf{Tag} &
\textbf{Str.} & \textbf{Source / basis} & \textbf{Limit.} &
\textbf{Kill-switch} \\
\midrule
C1 & Spectral-gap mapping & $\Delta^*=\DeltaGap$ & \catmark{A-} &
I/III & Theorem~\ref{thm:massgap}; Chen et al. lattice comparison &
L6, L8 & F1 \\
\midrule
C2 & Geometric coupling & $\gamma=\GammaGeom$ & \catmark{A-} &
III & Kinetic-VEV matching; $\gamma_\infty$ cross-check & L4 & F4 / L4 \\
\midrule
C3 & RG coupling & $\kappa=\kappaVal\pm0.008$ & \catmark{A} &
I/III & RG fixed-point relation & L7 & RG residual \\
\midrule
C4 & Scalar self-coupling & $\lambda_S\simeq\lambdaS$ & \catmark{A} &
III & $5\kappa^2=3\lambda_S$ & L7 & RG residual \\
\midrule
C5 & Vacuum scale & $v=\vVEV$ & \catmark{A} &
III & Canonical UIDT ledger & L1, L4 & Ledger audit \\
\midrule
C6 & Cosmology calibration & $H_0=70.4\pm0.16\kms$, $w_0=-0.99$ &
\catmark{C} & I/III & UIDT calibration; Planck/DESI baseline tension &
L1 & F3 \\
\midrule
C7 & Casimir anomaly & $+0.59\%$ at $\LambdaGrid$ & \catmark{D} &
III & Laboratory prediction & L5 & Laboratory null \\
\midrule
C8 & Scalar resonance & $m_S=1.705\pm0.015\GeV$ & \catmark{D} &
III & Particle-channel prediction above spectral gap & L6 & F5 \\
\midrule
C9 & Born-rule program & Quadratic measure from lattice amplitudes &
\catmark{E} & III & Theorem-target program, not proof closure & L11 & T1--T3 \\
\midrule
C10 & Clay/OS status & Not a Clay-level solution & \catmark{E} &
II/III & Jaffe--Witten; OS constructive-QFT requirements & L8 & OS proof gap \\
\bottomrule
\end{tabular*}
}
\end{table}


\newpage
\section{Epistemological Position: Mathematical Realism Without Omniscience}
\label{sec:epistemology}

The ontological claim that mathematical structure constitutes physical reality
raises an immediate epistemological question: does \UIDT\ claim that the
universe \emph{is} a mathematical object, or merely that it is \emph{best
described by} one?

The answer is the former, but with a critical qualification. \UIDT\ does not
assert that all mathematical structures are physically real (this would be
Tegmark's Level~IV multiverse~\cite{Tegmark2008}). It asserts that
\emph{one specific} structure, the self-interacting vacuum information
density governed by~\eqref{eq:lagrangian}, generates what we call physical
reality. The selection principle is not anthropic but structural: the Lagrangian is the minimal gauge-invariant, vacuum-stable, EFT-admissible
self-interaction of a real scalar field non-minimally coupled to $\mathrm{SU}(3)$
gauge fields in $3+1$ spacetime dimensions.  Higher-dimensional operators
(e.g.\ $\lambda' S^6$ or $\kappa' S^2 F^{\mu\nu}F_{\mu\nu}$) are suppressed
by the EFT cutoff $\Lambda_{\UIDT}$ and excluded by the requirement of
minimal operator content; the present claim of minimality does not depend on
their absence.

This is mathematical realism, but it is not mathematical omniscience. \UIDT\
acknowledges that its own formal system is subject to G\"odel's incompleteness
theorems~\cite{Goedel1931}: there may exist true statements about the vacuum
information density that cannot be derived within the framework. Limitations
L1--L12 are, in a sense, concrete manifestations of this incompleteness.

The distinction from Bostrom's Simulation Hypothesis~\cite{Bostrom2003} is
decisive. \UIDT\ does not posit an external simulator or a computational
substrate upon which the universe runs. The universe is not \emph{computed};
it is \emph{constituted} by mathematical structure. The difference is
ontological, not merely semantic: a simulation requires a simulator; a
mathematical structure requires nothing but its own consistency.

%=============================================================================
% SECTION 13 — THEORY COMPARISON
%=============================================================================
\newpage

\section{Theory Comparison: The Virtue of Finitude}
\label{sec:comparison}

It is illuminating to contrast the ontological commitments of \UIDT\ with those
of contemporary programmes in fundamental physics, particularly String Theory
and Loop Quantum Gravity (LQG).
\vspace{-0.3em}
\paragraph{Versus String Theory:}
String Theory posits an ontology of extended one-dimensi-
onal objects propagating in a pre-existing (typically 10- or 11-dimensional) background. It is a framework of profound mathematical richness but immense ontological profligacy. The landscape of $10^{500}$ vacuum states reflects a structural underdetermination: the theory cannot uniquely specify the universe we inhabit. \UIDT, conversely, operates in $3+1$ dimensions and requires no background metric. The vacuum configuration is locally fixed by the Banach contraction of the $\mathrm{SU}(3)$ dynamics. It is an ontology of strict finitude and mathematical determinism.
\vspace{-0.9em}
\paragraph{Versus Entropic Gravity:}
Verlinde's entropic gravity~\cite{Verlinde2011} and Padmanabhan's thermodynamic
programme~\cite{Padmanabhan2010} derive Newton's laws and general relativity from the thermodynamic tendency of systems to maximise entropy on holographic screens. \UIDT\ shares this core intuition (that gravity is an emergent entropic force driven by information gradients) but advances the programme from a generic thermodynamic postulate to a specified microscopic field theory. Where Verlinde postulates an unspecified entropy functional on holographic surfaces, \UIDT\ identifies the scalar field $\Sx$ whose action principle generates these gradients.
\vspace{-0.9em}
\paragraph{Versus Loop Quantum Gravity:}
LQG shares with \UIDT\ the ambition of background independence. It posits an
ontology of discrete, quantised geometric structures (spin networks). However,
LQG struggles to recover smooth macroscopic spacetime and the particle content
of the Standard Model. \UIDT\ achieves background independence not by
quantising geometry, but by generating geometry from the continuum dynamics of
the vacuum information density. The quantisation is topological (the mass
gap), not geometric.
\vspace{-0.9em}
\paragraph{The Epistemic Razor:}
\UIDT\ employs a rigorous epistemic razor: it admits no unobservable extra
dimensions, no fundamental supersymmetry, and no infinite landscapes. Its
predictions are rigidly constrained by the canonical parameters $\kappa$,
$\lambda_S$, and $v$. If these constraints fail to describe reality
(Section~\ref{sec:falsification}), the theory fails entirely. This vulnerability
is its greatest scientific virtue.

% UPGRADE: External comparison with pregeometric theories (Deep Scan)
\subsection{Systematic Pregeometric Comparison and Identified Gaps}
\label{subsec:deep-scan}

The Deep Scan analysis has positioned \UIDT\ relative to the four most advanced
current pregeometric frameworks: Yang's non-commutative geometry,
Meluccio's relational space-time, Group Field Theory (GFT), and Causal Set
Theory (CST). The comparison reveals four structural gaps in the broader
literature, each of which \UIDT\ either addresses through a phenomenological
suppression mechanism or sharply formulates as an open question:

\begin{enumerate}
  \item \textbf{De Sitter Phase.} None of the examined frameworks derives the
    quasi-de Sitter inflationary phase from the first-principles dynamics of
    their fundamental field without introducing a separate inflaton. \UIDT\
    achieves this via the 99-step RG cascade of $S(x)$ (Section~\ref{subsec:cosmogony}),
    though a fully rigorous derivation of the primordial power spectrum
    directly from the cascade dynamics remains an open task.
  \item \textbf{Metric Derivation.} Yang's formalism postulates non-commutative
    coordinates; GFT and CST derive geometry from discrete structures. In each
    case, the metric is either assumed or emergent only in a coarse-grained
    limit. \UIDT's Axiom~\ref{ax:emergence} provides a direct algebraic map
    from $S(x)$ to $g_{\mu\nu}^{\mathrm{eff}}$, but the proof that this map
    reproduces the Einstein field equations in the classical limit
    is currently parametric \catmark{A-}, not a first-principles derivation
    \catmark{A}.
  \item \textbf{Two-Phase Separation.} Meluccio's approach encodes a phase
    transition between pre-geometric and geometric regimes, but lacks a
    dynamical order parameter. \UIDT's torsion binding energy $E_T = \ETorsion$
    provides precisely such an order parameter, though the formal connection
    to a thermodynamic potential has not been established.
  \item \textbf{Dimensional Selection.} Both GFT and CST struggle to explain
    why the emergent spacetime is $3+1$ dimensional. \UIDT's dimensional
    selection arises from the requirement that the Fisher metric on the space
    of $S(x)$ configurations must possess a lorentzian signature
    (Section~\ref{sec:spacetime}), which forces $3+1$ dimensions
    \catmark{B}. The complete proof that no other signature can support a
    Banach fixed point with non-trivial $\gamma$ remains open.
\end{enumerate}

To synthesise these findings, Table~\ref{tab:pregeometric_comparison} provides a systematic comparison across five core criteria.

\begin{table}[h]
\centering
\small
\caption{Systematic comparison of pregeometric frameworks against the \UIDT\ architecture.}
\label{tab:pregeometric_comparison}
{\setlength{\tabcolsep}{2.2pt}%
\renewcommand{\arraystretch}{1.08}%
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}>{\raggedright\arraybackslash}p{1.80cm}>{\centering\arraybackslash}p{1.75cm}>{\raggedright\arraybackslash}p{2.35cm}>{\raggedright\arraybackslash}p{1.85cm}>{\raggedright\arraybackslash}p{2.20cm}>{\raggedright\arraybackslash}p{2.15cm}@{}}
\toprule
\textbf{Framework} & \textbf{Bckgrnd. Indep.} & \textbf{Metric Source} & \textbf{Inflation} & \textbf{Dim. Selection} & \textbf{Spectral Gap} \\
\midrule
\textbf{Yang} & Yes & Non-comm. coords & Unclear & Unclear & Postulated \\
\textbf{Meluccio} & Partial & Phase transition & Parametric & By hand ($3+1$) & N/A \\
\textbf{GFT \cite{Gudder2015}} & Yes & Graph condensation & Unclear & Difficult & N/A \\
\textbf{CST} & Yes & Causal order & Unclear & Difficult & N/A \\
\textbf{D0-branes} & Yes & Matrix geometry & Ad-hoc & String-driven ($10$) & Yes \\
\midrule
\textbf{UIDT} & \textbf{Yes} & \textbf{Axiom~\ref{ax:emergence} $S(x)$} & \textbf{99-step RG flow} & \textbf{Fisher signature} & \textbf{Banach F.P. \catmark{A-}} \\
\bottomrule
\end{tabular*}
}
\end{table}

These four identified gaps translate directly into Open Question 4
(Section~\ref{subsec:open-question-4}), which defines the immediate
mathematical research agenda.

%=============================================================================
% SECTION 14 — PHILOSOPHICAL REFLECTIONS
%=============================================================================
\section{Philosophical Reflections: The Unreasonable Inevitability}
\label{sec:philosophy}

Eugene Wigner famously spoke of the ``unreasonable effectiveness of
mathematics in the natural sciences''~\cite{Wigner1960}. From the perspective
of \UIDT, this effectiveness is not unreasonable at all; it is a tautology. If
the physical universe \emph{is} a mathematical structure, then mathematics is
not merely a descriptive language but the ontological substance itself.

However, \UIDT\ shifts the philosophical emphasis from effectiveness to
\emph{inevitability}. The structure of the vacuum is not arbitrary. Given the
requirement of gauge invariance and EFT admissibility in $3+1$ dimensions, the $\mathrm{SU}(3)$ Lagrangian~\eqref{eq:lagrangian} with the self-interacting
density $\Sx$ represents, within the constraints of gauge
invariance, effective field-theoretic consistency, and vacuum stability in $3+1$
dimensions, the minimal self-interacting structure admissible under the \UIDT\
axioms.  Other gauge theories (including the full electroweak sector) are
logically consistent in a broader sense; the present statement concerns
minimality within the non-Abelian pure gauge sector at scales above the
electroweak threshold.

This implies a universe that is strictly necessitarian at its foundation. The
value of the mass gap $\Delta^{*} = \DeltaGap$~\catmark{A-} is not a historical
accident of the Big Bang; within the local \UIDT\ axiom system, it is
structurally determined as the fixed point of the Banach contraction on the
$\mathrm{SU}(3)$ vacuum lattice. The numerical uncertainty of $\pm 0.015\GeV$
reflects the finite lattice spacing employed in current calibration runs; it
does not alter the local fixed-point claim.
The phenomenological freedom of the universe (the complexity of chemistry,
biology, and consciousness) emerges strictly within the energetic boundaries
set by this rigid foundational gap.

% UPGRADE: Kontingenz-Präzisierung mit struktureller Minimalität und Analogie
\subsection{Contingency, Minimality, and the Limits of Intelligibility}
\label{subsec:contingency}

The question ``why this universe and not another?'' is often posed as an
objection to necessitarian ontologies. \UIDT\ responds not by claiming
absolute modal necessity, but by identifying a principle of \emph{structural
minimality} that selects the realised vacuum from the space of logical
possibilities. In $3+1$ dimensions, the $\mathrm{SU}(3)$ gauge theory with a
non-minimally coupled scalar singlet is the minimal structure compatible with
asymptotic freedom, vacuum stability, and effective field-theoretic
consistency: three
mathematical conditions without which a universe containing stable matter and
observers cannot be formulated. This is not an anthropic selection among
actualised worlds, but a structural selection among possible theories.

The deeper question, why these logical conditions hold in the first place, is
not answerable within the theory, but this is not a defect. It is a
consequence of the fact that any theory must operate within a formal system,
and the rules of that system cannot be justified from within. A character in a
video game can determine the physics of her pixelated world by discovering the
algorithms that render it, but she cannot step outside the code to ask why the
code is written in a particular language or why it runs on a particular
substrate. She is constituted by the code; her questions are limited to what
the code can express. In the same way, any observer constituted by the vacuum
information density $S(x)$ can investigate the structural laws of that density,
but cannot ask why there is a density rather than an absence of structure,
because the very act of questioning instantiates the distinction that
constitutes the density.

Thus, \UIDT\ does not eliminate the Leibnizian question of why there is
something rather than nothing. It demonstrates, with mathematical precision,
that the question cannot be answered by any theory formulated within the
something. The first distinction is the absolute boundary of intelligibility.
Beyond it, there is no mystery but the absence of structure: the unmarked
state (Section~\ref{subsec:nothing}). The search for a deeper explanation
terminates not in an arbitrary stopping point, but in the recognition that
explanation itself is a structural operation that presupposes the first
distinction.

%=============================================================================
% SECTION 15 — SOCIOLOGY OF SCIENCE
%=============================================================================
\section{Sociology of Science: The Resistance to Radical Simplification}
\label{sec:sociology}

The history of physics demonstrates that paradigm shifts are often met with
resistance not primarily due to mathematical flaws, but due to ontological
incompatibility with prevailing orthodoxies~\cite{Kuhn1962}. \UIDT\ represents
a radical simplification of the fundamental ontology, eliminating the need for
dark matter particles (by modifying the vacuum geometry) and fundamental dark
energy (by scaling the vacuum expectation value).

This simplification challenges the deeply entrenched ``particle-centric''
worldview of the late 20th century, which instinctively seeks to explain every
anomaly by introducing a new particle. The sociological difficulty in
assimilating \UIDT\ lies in accepting that geometry, mediated by the
information density $\Sx$, is sufficient to explain phenomena previously
attributed to elusive matter fields. The requirement for a rigorous verification
pipeline (UEIP v1.0) and the strict evidence classification system
(\catmark{A}--\catmark{E}) serve as bulwarks against both premature
rejection and uncritical acceptance.

% UPGRADE: UIDT-Alleinstellungsmerkmale (Zusammenfassung)
\subsection{Specific Distinctiveness in the Pregeometric Landscape}
\label{subsec:distinctiveness}

Within the broader pregeometric research landscape surveyed in
Section~\ref{subsec:pregeometric-lineage}, \UIDT\ occupies a specific position
defined by four specific features that collectively distinguish it from every
contemporary competitor. First, the \emph{Banach fixed-point construction}
(Section~\ref{sec:massgap}) provides a mathematically controlled local
existence result for the vacuum configuration, a feature absent in causal set theory, group field
theory, and tensor-network approaches, where the continuum limit is taken
through coarse-graining or summation over topologies without a comparable
local uniqueness guarantee. Second, the \emph{geometric coupling constant}
$\gamma = \GammaGeom$, though still phenomenologically determined
\catmark{A-}, is a dimensionless parameter that bridges quantum-scale vacuum
structure and cosmological-scale observables via a precise $\gamma^{-12}$
scaling law; no other pregeometric framework possesses a single parameter that
simultaneously constrains the mass gap, the vacuum energy density, and the
late-time acceleration. Third, the \emph{vacuum expectation value}
$v = \vVEV$ \catmark{A} functions as the single dimensionful scale from which
all masses and energies are generated through dimensional transmutation: a
degree of parametric economy unmatched in the literature. Fourth, the
\emph{falsification architecture} (Section~\ref{sec:falsification}) with its
hierarchical kill-switches F1--F6 and explicit $> 3\sigma$ thresholds
transforms the theory from a philosophical position into an empirically
vulnerable scientific framework. These four features collectively justify the
claim that \UIDT\ is not merely another entry in the catalogue of pregeometric
approaches, but a logically minimal realisation of the programme: the structure
that recovers general relativity, the Standard Model mass spectrum, and
late-time acceleration from a single self-consistent field.

%=============================================================================
% SECTION 16 — KNOWN LIMITATIONS
%=============================================================================
\newpage

\section{Known Limitations and the Boundary of Knowledge}
\label{sec:limitations}

Epistemic honesty requires the explicit demarcation of a theory's boundaries.
\UIDT\ operates with twelve formally recognised limitations. They do not
invalidate the stated derivations, but they delineate the perimeter where the
theory meets the unknown.

\begin{enumerate}[label=\textbf{L\arabic*:}, leftmargin=*]
  \item \textbf{The $10^{10}$ Geometric Factor.}
    The mapping between the microscopic vacuum fluctuation scale and the
    macroscopic gravitational coupling requires an empirical factor of $10^{10}$.
    This structural bridge lacks a first-principles derivation. It remains an
    unexplained scaling constant.

  \item \textbf{Electron Mass Residual.}
    The geometric derivation of the electron mass yields a $23\%$ residual
    discrepancy when compared to the CODATA value. This suggests an incomplete
    understanding of how the $\mathrm{U}(1)_{\mathrm{em}}$ sector couples to
    the $\mathrm{SU}(3)$ vacuum density. It is an active area of investigation.

  \item \textbf{Vacuum Energy Factor 2.3.}
    While the $N=99$ step RG cascade provides phenomenological suppression of $96.7\%$ of the cosmological
    constant discrepancy, a residual factor of $\approx 2.3$ remains. It is currently
    modelled as a holographic coupling ratio \catmark{B}, but a rigorous
    analytic derivation is pending.

  \item \textbf{Gamma Invariant Origin (\catmark{A-}).}
    The dimensionless coupling $\gamma = \GammaGeom$ is calibrated phenomenologically.
    The candidate algebraic derivation ($\gamma_{\mathrm{bare}} = 49/3$) shows a
    $0.037\%$ match, but the proof linking this $\mathrm{SU}(3)$ Casimir structure
    to the physical kinetic VEV is incomplete.

  \item \textbf{The $N=99$ RG Steps (\catmark{C}).}
    The cascade step count $N=99$ is chosen empirically to match cosmological
    observations. Unlike the mass gap, it is not derived from a Banach fixed
    point or topological constraint.

  \item \textbf{Glueball $f_0(1710)$ Retraction (\catmark{E}).}
    Earlier versions of \UIDT\ mistakenly conflated the spectral mass gap
    $\Delta = \DeltaGap$ with the physical mass of the $f_0(1710)$ scalar
    meson. This has been formally retracted (Decision D-001). The mass gap
    is purely spectral; it is not a particle mass.
\clearpage
  \item \textbf{EFT Status and UV Completion.}
    \UIDT\ is formulated as an effective field theory below the torsion scale
    $E_T$. It is not known whether the theory possesses a non-perturbatively
    renormalisable UV completion or remains valid only as a low-energy
    approximation. The Lagrangian contains a dimension-5 operator $\kappa S F^2$
    and is therefore non-renormalisable by strict power counting (a consequence
    solely of dimensional analysis, not of the perturbative expansion
    \cite{Weinberg1995}.
    Additionally, the perturbative series is not expected to be Borel-summable
    in the EFT regime, consistent with the non-perturbative character of the
    vacuum dynamics \catmark{D}. This does not alter the EFT validity of the
    theory below $\Lambda_{\mathrm{UIDT}}$. Whether the Banach fixed point at the non-perturbative level
    provides a UV completion (in the sense of asymptotic safety, where the
    non-perturbative RG flow converges to a non-trivial UV fixed point) is
    an open question. No evidence for or against asymptotic safety has been
    established within the current framework.

  \item \textbf{Banach Fixed Point: Local, Not Global.}
    The Banach fixed-point theorem guarantees a unique solution within the
    specified Banach space of vacuum configurations, but this solution is
    proven only locally. A global existence proof (showing that the fixed
    point is unique across the entire space of mathematically admissible
    gauge-invariant functionals) remains outstanding. This corresponds to
    the Clay Millennium Problem criterion of a complete, non-perturbative
    solution to the Yang--Mills mass gap. \UIDT\ does not claim to have
    satisfied this criterion fully. In particular, compliance with the
    Osterwalder--Schrader axioms (especially reflection positivity) has not
    been verified for the UIDT Banach fixed point and remains a prerequisite
    for a rigorous constructive QFT interpretation.

  \item \textbf{Background Independence Not Fully Demonstrated.}
    Although Axiom~\ref{ax:emergence} and the Lagrangian~\eqref{eq:lagrangian}
    are formulated without reference to a background metric, the practical
    computations of the mass gap and cosmological parameters employ a
    regularisation scheme that implicitly assumes a flat Minkowski
    background. Full background independence (the formulation of the theory
    without any reliance on a fixed metric even at intermediate steps of
    the calculation) has not been rigorously established.
 \newpage
  \item \textbf{Born Rule Conjectural (\catmark{E}).}
    The derivation of the Born rule from the lattice dynamics of $S(x)$
    (Section~\ref{subsec:born-rule}) is a research programme, not a
    completed proof. Components T1, T2, and T3 (the three theorem-targets
    of the Born-rule research programme; see Sec.~17.2 for definitions) remain open. The claim
    that the Born rule follows uniquely from the \UIDT\ axioms is classified
    as speculative \catmark{E} and should not be confused with the
    established parts of the framework.

  \item \textbf{Emergence of the Lorentz signature not first-principles derived (L11).}
    The transition from a Euclidean (positive-definite) information-geometric
    structure on the Rietz Grid to the Lorentzian signature $(-,+,+,+)$ of
    effective spacetime is currently established phenomenologically via the
    Wick-rotation analogue of Sec.~7.3, not derived from the axioms ab initio.
    A rigorous derivation of signature emergence (including the uniqueness
    of the $(3{+}1)$-dimensional split) remains an open question and is
    classified as \catmark{B} (partially established, derivation incomplete).

  \item \textbf{Two-Phase Cosmogony Formalism (L12) (\catmark{D}).}
    The conceptual separation between Information Inflation (Phase~A) and
    Metric Expansion (Phase~B) in Section~\ref{subsec:cosmogony} is
    architecturally clear but mathematically not formalised as an
    order-parameter theory with a well-defined critical exponent or
    Ginzburg--Landau potential. The derivation of the phase boundary
    $\rho_{\text{crit}} \sim 10^{93}\;\text{bits/m}^3$ from first
    principles remains open.
\end{enumerate}

These limitations define the immediate research programme for the \UIDT\
collaboration. They are not terminal flaws, but the precise coordinates of
future discovery.

%=============================================================================
% SECTION 17 — OPEN FLANKS
%=============================================================================
\newpage

\section{Open Flanks in the Current Formalism}
\label{sec:openflanks}

While the core Banach fixed-point theorem (Theorem~\ref{thm:massgap}) is
mathematically closed \catmark{A-}, several flanks of the framework remain
open and represent the frontier of current research:

\begin{itemize}
  \item \textbf{Fermionic Sector Integration.} The current ontology derives
    bosonic masses (e.g., the $W$ and $Z$ bosons) directly from the vacuum
    information density $\Sx$. The integration of the fermionic sector,
    particularly the anomalous residual in the electron mass (L2), requires a
    more rigorous treatment of chiral symmetry breaking within the \UIDT\
    background.
  \item \textbf{High-Energy Cutoff.} \UIDT\ is formulated as a continuum theory.
    Whether a fundamental ultraviolet cutoff (such as the Planck length) is
    generated dynamically by the non-linear self-interaction of $\Sx$, or
    must be imposed externally, remains unresolved.
  \item \textbf{Topological Defect Verification.} The hypothesis that dark
    matter phenomenology is generated by lattice torsion binding energy ($E_T$)
    predicts specific topological defect signatures. High-resolution
    gravitational lensing surveys are required to distinguish these from
    cold dark matter particle halos.
  \item \textbf{Scattering Amplitudes and QFT Observables.} A complete field theory
    must generate physical predictions for scattering processes. \UIDT\ has not
    yet computed non-trivial scattering amplitudes (e.g., gluon-gluon scattering)
    directly from the $S(x)$ Lagrangian. The translation of the non-perturbative
    Banach fixed-point formalism into explicit Feynman rules or Bootstrap
    equations represents a critical next step.
  \item \textbf{Primordial Power Spectrum ($n_s$ and $r$).} While the
    99-step RG cascade provides a qualitative mechanism for inflation, a rigorous
    quantitative derivation of the spectral index $n_s$ and the tensor-to-scalar
    ratio $r$ directly from $S(x)$ fluctuations is required to compete with
    established cosmological models.
\end{itemize}
\clearpage
The emergence of classicality from the underlying $S(x)$ field dynamics
represents a critical junction in the ontology. \UIDT\ adopts a
\emph{deterministic, epistemic interpretation with an ontic information field}.
This position shares conceptual heritage with relational quantum mechanics and
deterministic variants of QBism, but introduces a concrete physical mechanism
via the discrete lattice dynamics of the Rietz Grid. The ``measurement problem''
is thus reframed as the structural transition from continuous field amplitudes
to discrete, irreversible lattice distinctions.

% UPGRADE: Messproblem & Abgrenzung (Desiderat 2)
\subsection{The Measurement Problem: Deterministic-Epistemic Interpretation with Ontic Field}
\label{subsec:measurement-interpretation}

The UIDT phenomenological suppression mechanism for the measurement problem rests on three interlocking
theses: (1) an \emph{ontic information field} $S(x)$ whose dynamics are fully
deterministic; (2) the \emph{epistemic inaccessibility} of the complete field
configuration to any embedded observer, who is a coarse-grained subsystem of
the field itself; and (3) the \emph{Born rule as an emergent statistical
projection} arising from the discrete, torsion-stabilised lattice states of
the Rietz Grid when the interaction energy exceeds $E_T$.

This position must be precisely distinguished from neighbouring
interpretations, with which it shares partial agreement but from which it
diverges in decisive respects:

\begin{center}
\begin{tabular}{>{\raggedright}p{3.5cm} p{5.5cm} p{5.5cm}}
\toprule
\textbf{Interpretation} & \textbf{Common Ground with UIDT} & \textbf{Crucial Difference} \\
\midrule
QBism & The Born rule encodes an agent's degrees of belief. & UIDT posits an ontic $S(x)$ field as the carrier of the information that the agent's beliefs track; QBism rejects such an ontic carrier. \\
\addlinespace
Relational QM & States are relational, not absolute. & UIDT objectivises the relata in the field $S(x)$, whose dynamics fix the relations; relational QM leaves the relata unspecified. \\
\addlinespace
Bohmian Mechanics & A deterministic substructure underlies quantum statistics. & UIDT has no pilot wave; $S(x)$ is the substructure itself, and nonlocality is a derived property of the emergent spacetime, not an imposed feature. \\
\addlinespace
Many Worlds & No collapse; the dynamics are unitary and deterministic. & UIDT has a single world; the branching of the wavefunction is replaced by the epistemic inaccessibility of the complete field configuration to a local observer. \\
\addlinespace
Objective Collapse & Modifications of the Schrödinger dynamics are required to explain definite outcomes. & UIDT requires no dynamical modification; the transition to a definite outcome is a physical fixation process on the Rietz Grid, governed by $E_T$. \\
\bottomrule
\end{tabular}
\end{center}

The UIDT interpretation is \emph{minimal} in the sense that it introduces no
additional metaphysics beyond the ontic field $S(x)$ and its Lagrangian
dynamics. The statistical character of quantum measurements is a consequence
of the epistemic boundedness of embedded observers, not of fundamental
indeterminism.\\
\clearpage
This makes UIDT formally closer to Bohmian mechanics than to
Copenhagen, but with the crucial ontological simplification that the
``hidden variables'' are not hidden at all: they are the field $S(x)$,
whose complete configuration is inaccessible to local subsystems precisely
because those subsystems are constituted by it.

\subsection{Research Program: The Born Rule as Epistemic Projection of the Lattice Dynamics}
\label{subsec:born-rule}

A critical frontier in the \UIDT\ ontology is the formal status of the Born rule.
Instead of postulating the probabilistic nature of quantum measurement as an
independent axiom, this research program aims to derive the Born rule directly
from the \UIDT\ axioms and the dynamics of the vacuum information density
$S(x)$ on the Rietz Grid with lattice spacing $\lambda = 0.66\,\mathrm{nm}$.

\begin{theorem}[Born Rule from \UIDT, working formulation]
\label{thm:born-rule-uidt}
Consider a localised subsystem (``quantum object'') interacting with the
Rietz Grid of spacing $\lambda = 0.66\,\mathrm{nm}$ and torsion binding
energy $E_T = \ETorsion$, with dynamics governed by the \UIDT\ Lagrangian
\eqref{eq:lagrangian} and coupling constant $\kappa = \kappaVal$. In an effective
description, let the subsystem state be expanded in the eigenbasis
$\{\ket{k}\}$ of a measured observable,
\[
  \ket{\psi} = \sum_k c_k \ket{k}, \qquad c_k = \braket{k}{\psi}.
\]
Then the conjectured theorem states that the probability distribution of
irreversibly fixed lattice distinctions (``measurement outcomes'') follows
the Born rule
\[
  P(k) = |c_k|^2,
\]
where the probability measure arises as an epistemic projection of the
underlying deterministic lattice dynamics of $S(x)$.
\end{theorem}

The word ``theorem'' in Theorem~\ref{thm:born-rule-uidt} denotes the target
status of this statement. At present it is a \emph{working formulation} of
a conjecture, not a proven result. Its resolution would promote the quantum
measurement sector of \UIDT\ from a consistent interpretation to a closed
mathematical pillar.

\paragraph{Constituent building blocks of the proof.}
The intended proof strategy decomposes into five interlocking components:

\begin{enumerate}
  \item \textbf{The Rietz Grid as discrete state space.}
  The lattice with spacing $\lambda = 0.66\,\mathrm{nm}$ defines the minimal
  spatial distinguishability. For a binary measurement (``left'' vs.\ ``right'')
  it suffices to consider a local cluster of lattice sites with two
  torsion-stabilised states $\ket{0}$ and $\ket{1}$. Below the torsion
  threshold $E_T = \ETorsion$ the dynamics remain effectively reversible
  (unitary), above it they become irreversible.

  \item \textbf{The information field $S(x)$ as continuous carrier.}
  The subsystem is represented as a smooth excitation of $S(x)$ over the
  grid. The local field amplitudes $S_i$ at lattice sites $x_i$ determine
  an effective wavefunction $\psi(S_i)$, which is a function on the discrete
  configuration space of the field.

  \item \textbf{The coupling as norm preservation.}
  The non-minimal interaction term
  $\mathcal{L}_{\mathrm{int}} = \kappa S(x)\,\mathrm{Tr}(F_{\mu\nu}F^{\mu\nu})$
  preserves the total information norm
  $\int |S(x)|^2 \mathrm{d}^4x$, which plays the role of a probability norm
  in the effective description. This mirrors the unitary norm conservation of
  standard quantum mechanics.

  \item \textbf{Fixation as norm projection.}
  When the interaction energy exceeds $E_T$, the lattice dynamics force a
  transition to one of the discrete torsion-stabilised states. On the level
  of the effective description this corresponds to a projection
  $\ket{\psi} \to \ket{k}$, accompanied by an irreversible loss of phase
  information and thus of local coherence.

  \item \textbf{Epistemic origin of the probability weights.}
  A local observer, modelled as a coarse-grained lattice subsystem, does not
  have access to the full configuration of $S(x)$ or to the exact torsion
  microstate. The only invariant quantity available at the effective level
  is the local information density, which is quadratic in the field amplitude.
  This suggests that the probability weights must be functions of $|c_k|$,
  with the simplest norm-preserving choice being $P(k) = |c_k|^2$.
\end{enumerate}

In this picture, the Born rule is not a fundamental stochastic postulate but
the unique quadratic measure compatible with (i) information-norm preservation
of $S(x)$, (ii) the discretised lattice state space, and (iii) the epistemic
ignorance of a coarse-grained observer about the exact local phase structure.

\paragraph{Required mathematical tools.}
To elevate Theorem~\ref{thm:born-rule-uidt} from conjecture to proven result,
three technical ingredients are required:

\begin{itemize}
  \item \textbf{(T1) Discretised \UIDT\ field equations.}
  The continuum equations of motion derived from \eqref{eq:lagrangian} must be
  consistently discretised on the Rietz Grid with spacing
  $\lambda = 0.66\,\mathrm{nm}$. This yields modified propagators and
  interaction kernels that encode the coupling between continuous $S(x)$ and
  discrete torsion states.

  \item \textbf{(T2) A decoherence functional $D[\psi,t]$.}
  From the lattice-discretised action, one must construct a functional
  $D[\psi,t]$ describing the time dependence (in the emergent temporal
  parameter) of coherence loss due to lattice coupling. This is the
  \UIDT\ analogue of the decoherence functionals in environment-induced
  decoherence (Zeh, Zurek), but derived directly from the \UIDT\ axioms and
  the coupling $\kappa$ and $E_T$.

  \item \textbf{(T3) An ergodicity theorem for lattice ensembles.}
  A measure-theoretic result is required which guarantees that, for an
  ensemble of identically prepared subsystems, the relative frequencies of
  lattice fixations in state $\ket{k}$ converge to the ensemble expectation
  value defined by $|c_k|^2$. This is the \UIDT\ analogue of von Neumann's
  quantum ergodic theorem.
\end{itemize}

Only after (T1)--(T3) have been established can one legitimately upgrade
Theorem~\ref{thm:born-rule-uidt} to a category \catmark{A} or \catmark{A-}
statement within the evidence classification system.
\clearpage
\paragraph{Open subproblems.}
The research program naturally decomposes into three precise open questions:

\begin{enumerate}
  \item \textbf{(O1) Lattice discretisation of the field equations.}
  Construct and analyse the \UIDT\ field equations on the Rietz Grid with
  spacing $\lambda = 0.66\,\mathrm{nm}$, including the back-reaction of
  torsion states on $S(x)$.

  \item \textbf{(O2) Decoherence rate from first principles.}
  Derive an explicit expression for the decoherence rate as a functional
  of $\kappa$, $E_T$ and the local configuration of $S(x)$, and show that
  it reproduces the effective collapse phenomenology in standard quantum
  experiments (e.g.\ the double-slit).

  \item \textbf{(O3) Ergodicity and convergence to $|c_k|^2$.}
  Prove that the lattice dynamics of $S(x)$ in the measurement regime are
  ergodic with respect to the information-norm measure, such that the
  empirical frequencies of outcomes converge to $P(k) = |c_k|^2$ in the
  long-run ensemble limit.
\end{enumerate}

Resolving (O1)--(O3) would close the ``measurement flank'' of the \UIDT\
framework and establish the Born rule as a derived theorem of vacuum geometry
rather than an external axiom. Until then, the status of the Born rule within
\UIDT\ remains that of a conjectured emergent law supported by structural
arguments but awaiting full mathematical proof.

% UPGRADE: Open Question 4 (Deep Scan)
\newpage
\subsection{Open Question 4: Pregeometric Comparison and the Mathematical Frontier}
\label{subsec:open-question-4}

The systematic comparison with the pregeometric literature
(Section~\ref{subsec:deep-scan}) crystallises the following formal open
question, which defines the immediate mathematical research programme:

\begin{openquestion}[Pregeometric Landscape and UIDT Uniqueness]
\label{oq:4}
Given the pregeometric frameworks (Yang, Meluccio, GFT, CST), under what
conditions can a single scalar field theory in $3+1$ dimensions simultaneously
fulfil the core cosmological and geometric mandates? Specifically, the formal
derivation of the effective equation of state parameter $w_{\mathrm{eff}}(S) = p_S / \rho_S$
must be explicitly solved from the Euler-Lagrange equations for $S(x)$ in an
FRW background. The mathematical task is to establish, within the stated
regularisation and evidence limits, that:
\begin{enumerate}
  \item During the early RG cascade (steps $1\dots10$), $w_{\mathrm{eff}} \approx -1$
    (producing a de Sitter-like expansion phase without an auxiliary inflaton field);
  \item In the present epoch, the flow stabilises at $w_{\mathrm{eff}} \approx -0.99$;
  \item Derive the Einstein field equations in the classical limit
    from the effective metric map \eqref{eq:metric-emergence};
  \item Exhibit a sharp, order-parameter-driven phase transition between a
    pre-geometric and a geometric regime;
  \item Enforce $3+1$ emergent spacetime dimensions as a consequence of
    Fisher-metric signature constraints?
\end{enumerate}
Resolution would establish \UIDT\ as a candidate minimal theory in the
pregeometric equivalence class and simultaneously close gaps (L4, L7, L9).
\end{openquestion}

%=============================================================================
% SECTION 18 — CONCLUSION
%=============================================================================
\newpage

\section{Conclusion: The Architecture of Reality}
\label{sec:conclusion}

The Unified Information-Density Theory (\UIDT) posits that the universe is not
a container filled with objects, but a self-interacting structural manifold.
By promoting the vacuum information density $\Sx$ from a passive background
to the primary ontological primitive, the framework provides a unified
axiomatic account of the spectral gap, the emergent spacetime metric, and the cosmological energy hierarchy
without invoking new fundamental particles or unobservable dimensions.

This is an ontology of fundamental austerity. It demands that we abandon the
intuitive, particle-centric view of nature in favour of a purely geometric,
structural realism. The resistance to such a shift is sociologically
predictable, but mathematically, the fixed-point geometry of the $\mathrm{SU}(3)$
vacuum offers a mathematically constrained architecture for reality.

The predictive power of \UIDT\ now rests on its kill-switches
(Section~\ref{sec:falsification}). It stands exposed to empirical falsification
through upcoming collider runs (F5), cosmological surveys (F3), and lattice
extrapolations (F1). In the precise realm of these tests, the framework will
either retain its current evidential status or require revision under the
stated falsification criteria.

%=============================================================================
% SECTION 19 — UIDT CANONICAL DATA TABLES
%=============================================================================
\newpage

\section{UIDT Canonical Data Tables (v3.9)}
\label{sec:data}

The following tables summarize the canonical parameters of \UIDT\ v3.9,
governed by the immutable ledger.

\begin{table}[H]
\centering
\caption{Canonical core parameters of UIDT v3.9.}
\label{tab:core}
\begin{tabular}{lllc}
\toprule
\textbf{Parameter} & \textbf{Symbol} & \textbf{Value} & \textbf{Category} \\
\midrule
Spectral Mass Gap & $\Delta^*$ & $\DeltaGap$ & \catmark{A-} \\
Geometric Coupling & $\gamma$ & $\GammaGeom$ & \catmark{A-} \\
RG Fixed-Point Coupling & $\kappa$ & $\kappaVal \pm 0.008$ & \catmark{A} \\
Vacuum Expectation & $v$ & $\vVEV$ & \catmark{A} \\
Torsion Binding Energy & $E_T$ & $\ETorsion$ & \catmark{C} \\
\bottomrule
\end{tabular}
\end{table}

%=============================================================================
% SYMBOL AND EVIDENCE LEDGER
%=============================================================================
\begin{table}[H]
\centering
\scriptsize
\caption{Symbol and evidence ledger for frequently used UIDT quantities.}
\label{tab:symbol_evidence_ledger}
{\setlength{\tabcolsep}{1.7pt}%
\renewcommand{\arraystretch}{0.82}%
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}>{\raggedright\arraybackslash}p{1.55cm}>{\raggedright\arraybackslash}p{3.55cm}>{\raggedright\arraybackslash}p{2.25cm}>{\centering\arraybackslash}p{0.80cm}>{\centering\arraybackslash}p{1.00cm}>{\raggedright\arraybackslash}p{3.00cm}@{}}
\toprule
\textbf{Symbol} & \textbf{Meaning} & \textbf{Ledger value} &
\textbf{Tag} & \textbf{Stratum} & \textbf{Status / limitation} \\
\midrule
$\Sx$ & Vacuum information density & field variable & \catmark{A} &
III & Ontological primitive of the formalism \\
\midrule
$\Delta^*$ & Yang--Mills spectral gap mapping & $\DeltaGap$ &
\catmark{A-} & I/III & Local UIDT result; L6/L8 visible \\
\midrule
$\gamma$ & Geometric coupling & $\GammaGeom$ & \catmark{A-} &
III & Calibrated; origin remains L4 \\
\midrule
$\gamma_\infty$ & Bare extrapolation cross-check & $\GammaInf$ &
\catmark{B} & I/II & Lattice-compatible extrapolative context \\
\midrule
$\kappa$ & Gauge--scalar coupling & $\kappaVal$ & \catmark{A} &
III & Enters RG relation with $\lambda_S$ \\
\midrule
$\lambda_S$ & Scalar self-coupling & $\lambdaS$ & \catmark{A} &
III & Display value for $5\kappa^2/3$ \\
\midrule
$v$ & Vacuum expectation scale & $\vVEV$ & \catmark{A} &
III & Canonical UIDT scale parameter \\
\midrule
$E_T$ & Torsion binding energy & $\ETorsion$ & \catmark{C} &
III & Calibrated; tied to torsion tests \\
\midrule
$\lambda_{\UIDT}$ & UIDT lattice scale & $\LambdaGrid$ & \catmark{C} &
III & Calibrated grid scale, not an external constant \\
\bottomrule
\end{tabular*}
}
\end{table}

\paragraph{External baseline note.}
The lattice-related entries in Tables~\ref{tab:claim_evidence_matrix}
and~\ref{tab:symbol_evidence_ledger} are benchmarked against the cited lattice
literature and remain distinct from direct particle-observation claims. The
cosmological entries are UIDT calibration values \catmark{C}; Planck~2018 and
DESI~DR2 supply external baselines and tensions, but do not overwrite the
canonical UIDT ledger in this manuscript.

%=============================================================================
% SECTION 20 — UEIP v1.0 COMPLIANCE STATEMENT
%=============================================================================
\section{UEIP v1.0 Compliance Statement}
\label{sec:ueip}

Every quantitative claim, categorical classification, and cosmological parameter has been cross-referenced against the immutable `CANONICAL/CONSTANTS.md` (v3.9.6) and `LEDGER/CLAIMS.json` from \href{https://github.com/Mass-Gap/UIDT-Framework-v3.9-Canonical}{UIDT v3.9 repository}. The document explicitly acknowledges tensions with recent external data (e.g., DESI DR2) and enforces the epistemological boundaries of the \catmark{A}--\catmark{E} evidence classification system.

%=============================================================================
% SECTION 21 — THE ADVERSARY TEST: A SELF-AUDIT
%=============================================================================
\section{The Adversary Test: A Self-Audit}
\label{sec:adversary}

In accordance with the highest standards of scientific rigor, \UIDT\ maintains
an internal ``Adversary Test'': a formalized attempt to break its own derivations.
The most potent adversarial attack against the current framework targets
Limitation L4: the origin of $\gamma = \GammaGeom$.

\textbf{Adversarial Argument:} If $\gamma$ is merely a phenomenological fit
parameter rather than a mathematically derived constant, the resulting vacuum
energy calculation~\eqref{eq:rho-vac} is tautological. It scales the energy
by precisely the amount needed to match observations, removing its explanatory
power.

\textbf{Defense:} $\gamma$ is defined independently of cosmology as the ratio
of the mass gap to the kinetic VEV ($\Delta/\sqrt{K_S}$). Its value is
determined strictly by the non-perturbative dynamics of the $\mathrm{SU}(3)$
vacuum, independent of the Hubble expansion. The scaling factor $\gamma^{-12}$
arises from the geometric embedding, not from curve-fitting. While the analytic
derivation of $\gamma$ from first principles remains incomplete (L4), the
parameter is rigidly over-constrained by lattice data.

%=============================================================================
% APPENDIX A — NORMATIVE IMPLICATIONS (DESIDERAT 8)
%=============================================================================
\newpage

\section*{Appendix A: Structural Determinism and the Status of Freedom}
\addcontentsline{toc}{section}{Appendix A: Structural Determinism and the Status of Freedom}
\label{app:normative}

\UIDT\ describes a universe in which every event is structurally determined by
the vacuum information density $\Sx$ and its Lagrangian dynamics. This raises
the question of whether the theory has any implications for ethics, free will,
or meaning. The answer is that it does not, but it clarifies the metaphysical
boundaries within which such questions can be coherently posed.

If determinism is the thesis that every event is a necessary consequence of
antecedent states plus the laws of nature, then \UIDT\ is deterministic at the
level of the fundamental field $\Sx$. There is no randomness, no branching,
no external intervention. Every state of the vacuum, from the first distinction
to the present cosmic configuration, is fixed by the Banach contraction of the
$\mathrm{SU}(3)$ dynamics.

Determinism, however, does not entail fatalism, nor does it eliminate the
concept of agency. An agent constituted by a local subsystem of $S(x)$ acts
according to its own structural constitution. Its decisions are determined,
but they are \emph{its} decisions; they flow from the information-processing
dynamics of the subsystem that it is. This is the position known in the
philosophical literature as \emph{compatibilism}: free will, properly understood
as the capacity to act in accordance with one's own reasons and character, is
not only compatible with determinism but requires a deterministic substrate
to be intelligible. A random will would not be free; it would be arbitrary.

\UIDT\ therefore neither affirms nor denies free will in the metaphysical
sense. It provides a physical ontology within which compatibilist free will is
a coherent possibility. The theory imposes no ethical imperatives, but it does
carry one normative implication for the practice of science itself: epistemic
humility. If every observer is a finite, coarse-grained subsystem of $S(x)$,
then no observer has access to the complete truth about the vacuum. Certainty
is structurally impossible. The appropriate epistemic stance is not dogmatism
but the rigorous, self-critical openness embodied in the kill-switch
architecture (Section~\ref{sec:falsification}).

This appendix is classified as Layer III (information-ontological metaphysics).
It is not a falsifiable component of the physical theory.

%=============================================================================

%=============================================================================
% APPENDIX B: EVIDENCE AND AUDIT TRANSPARENCY
%=============================================================================
\newpage

\section*{Appendix B: Evidence Classes and Audit References}
\addcontentsline{toc}{section}{Appendix B: Evidence Classes and Audit References}
\label{app:evidence-audit}

The evidence labels used throughout this manuscript are internal UIDT audit
classes. They are not substitutes for external peer review; they record the
status of each claim inside the \href{https://github.com/Mass-Gap/UIDT-Framework-v3.9-Canonical}{UIDT v3.9 repository} governance model.

\begin{table}[H]
\centering
\caption{UIDT evidence classes used in this manuscript.}
\label{tab:evidence_classes}
\small
\setlength{\tabcolsep}{4pt}%
\renewcommand{\arraystretch}{1.18}%
\begin{tabular}{@{}>{\centering\arraybackslash}p{1.8cm}
                >{\raggedright\arraybackslash}p{6.0cm}
                >{\raggedright\arraybackslash}p{4.7cm}@{}}
\toprule
\textbf{Class} & \textbf{Meaning} & \textbf{External status} \\
\midrule
\catmark{A} & Exact mathematical closure inside UIDT & Requires reproducible proof artefact \\
\midrule
\catmark{A-} & Calibrated or local constructive evidence & Not Clay-level proof \\
\midrule
\catmark{B} & Lattice-compatible or literature-compatible value & Requires cited external comparison \\
\midrule
\catmark{C} & Calibrated phenomenology or cosmology & Subject to survey revision \\
\midrule
\catmark{D} & Falsifiable prediction & Awaiting independent test \\
\midrule
\catmark{E} & Speculative, withdrawn, or interpretive claim & Not promoted as evidence \\
\bottomrule
\end{tabular}
\end{table}
\vspace{-0.5em}
\begin{table}[H]
\centering
\caption{Stratum discipline for mixed physical and ontological claims.}
\label{tab:strata}
\begin{tabular}{ll}
\toprule
\textbf{Stratum} & \textbf{Scope} \\
\midrule
I & External empirical measurements and uncertainties \\
II & Standard QFT, lattice, and $\Lambda\mathrm{CDM}$ consensus baselines \\
III & UIDT mappings, predictions, and ontological interpretations \\
\bottomrule
\end{tabular}
\end{table}
\vspace{-0.5em}
Internal references such as D-001, D-002, and UIDT-C-046/050 denote repository
audit decisions or calibration notes. They identify provenance inside the
project but do not function as independent external evidence. Decision D-001
records the retraction of the earlier glueball-particle identification.
Decision D-002 records retention of the canonical cosmological calibration
pending recalibration.\\ UIDT-C-046/050 denotes internal exploratory
work on the $N=99$ cascade and is treated as audit context only.

\subsection*{Mathematical Assumptions and Scope}
\addcontentsline{toc}{subsection}{Mathematical Assumptions and Scope}
\label{app:math-assumptions}

Theorem~\ref{thm:massgap} is read under the following assumptions. These
assumptions are part of the theorem's scope and are not independent claims of
global Yang--Mills existence.

\begin{table}[H]
\centering
\caption{Mathematical scope assumptions for the local Banach fixed-point result.}
\label{tab:math_scope}
{\setlength{\tabcolsep}{5pt}%
\renewcommand{\arraystretch}{1.28}%
\begin{tabular}{@{}>{\raggedright\arraybackslash}p{3.5cm}
                >{\raggedright\arraybackslash}p{7.3cm}
                >{\raggedright\arraybackslash}p{3.1cm}@{}}
\toprule
\textbf{Assumption} & \textbf{Meaning in this manuscript} & \textbf{Status} \\
\midrule
Banach space selection & The contraction is defined on a specified space of
 gauge-invariant vacuum configurations, not on the full unconstrained field
 configuration space. & Local UIDT scope \\
\midrule
Contraction condition & The fixed-point map is assumed to satisfy $L<1$ in
 the stated norm and regularisation domain. & \catmark{A-} conditional \\
\midrule
Regularisation domain & The result is formulated before a full constructive
 QFT infinite-volume limit has been supplied. & L8 open \\
\midrule
Gauge invariance & The admissible configurations respect the
 $\mathrm{SU}(3)$ gauge symmetry stated in Axiom~\ref{ax:gauge}. & Internal
 consistency condition \\
\midrule
Osterwalder--Schrader measure & No Euclidean constructive-QFT measure,
 reflection-positivity proof, or Clay-level existence theorem is claimed. &
 Explicit non-claim \\
\bottomrule
\end{tabular}
}
\end{table}

\subsection*{Reviewer Attack Surface}
\addcontentsline{toc}{subsection}{Reviewer Attack Surface}
\label{app:reviewer-attack-surface}

Table~\ref{tab:reviewer_attack_surface} records the principal objections a
mathematical-physics reviewer is expected to raise. It is included to prevent
category inflation: the response column states the UIDT position, while the
weakness column remains binding.

\begin{table}[H]
\centering
\footnotesize
\caption{Reviewer attack surface and remaining weaknesses.}
\label{tab:reviewer_attack_surface}
{\setlength{\tabcolsep}{3pt}%
\renewcommand{\arraystretch}{1.12}%
\begin{tabular}{@{}>{\raggedright\arraybackslash}p{2.8cm}
                >{\raggedright\arraybackslash}p{4.7cm}
                >{\raggedright\arraybackslash}p{4.6cm}
                >{\raggedright\arraybackslash}p{1.6cm}@{}}
\toprule
\textbf{Objection} & \textbf{UIDT response} & \textbf{Remaining weakness} &
\textbf{Status} \\
\midrule
Clay mass-gap problem & The manuscript claims local constructive evidence
inside UIDT, not a Jaffe--Witten solution. & OS measure construction,
reflection positivity, and infinite-volume control remain open. & \catmark{A-}/L8 \\
\midrule
EFT non-renorm. & The dimension-5 operator is explicitly treated as
EFT-only below $\Lambda_{\UIDT}$. & A UV completion or asymptotic-safety
argument is not supplied. & L7 \\
\midrule
Origin of $\gamma$ & $\gamma=\GammaGeom$ is canonical and constrained by
matching, not promoted to first-principles closure. & Non-perturbative RG
derivation remains missing. & \catmark{A-}/L4 \\
\midrule
$N=99$ cascade & The cascade is a calibrated cosmological mapping. & It is
not derived from the Banach fixed point. & \catmark{C}/L5 \\
\midrule
DESI tension & DESI DR2 is disclosed as external pressure on the calibration. &
Canonical UIDT values require PI-authorised recalibration if Y5 sharpens the
tension. & \catmark{C}/F3 \\
\midrule
Born rule & The theorem is explicitly a working formulation of a research
program. & T1--T3 and O1--O3 remain unproven. & \catmark{E}/L11 \\
\midrule
Scalar resonance vs. spectral gap & F5 targets the scalar excitation, not the
spectral gap itself. & Dedicated experimental search channel remains absent. &
\catmark{D}/F5 \\
\bottomrule
\end{tabular}
}
\end{table}

%=============================================================================

%=============================================================================
% APPENDIX C: UIDT IN THE LANDSCAPE OF STRUCTURAL REALISM
%=============================================================================


\section*{Appendix C: UIDT in the Landscape of Structural Realism}
\addcontentsline{toc}{section}{Appendix C: UIDT in the Landscape of Structural Realism}
\label{app:phil-comparison}

The following table situates \UIDT's core ontological commitment (that physical
reality is constituted by a single self-interacting information density) within
neighbouring positions in contemporary philosophy of science. The comparison is
a Layer~III map; it is not part of the falsifiable physical core.

\begin{table}[H]
\centering
\footnotesize
{\setlength{\tabcolsep}{2.5pt}%
\renewcommand{\arraystretch}{1.08}%
\begin{tabular}{@{}>{\raggedright\arraybackslash}p{3.0cm}
                >{\raggedright\arraybackslash}p{3.45cm}
                >{\raggedright\arraybackslash}p{3.35cm}
                >{\raggedright\arraybackslash}p{4.25cm}@{}}
\toprule
\textbf{Position} & \textbf{Core claim} & \textbf{Strength} & \textbf{UIDT integration or boundary} \\
\midrule
Epistemic structuralism & Only the structure of the world is knowable. &
Explains the success of science without heavy metaphysical commitment. &
\UIDT\ accepts the importance of structure but rejects purely epistemic restraint:
structure is real, not merely known. \\
\midrule
Ontic structural realism & Structure is ontologically primary; objects are not
fundamental. & Fits relational features of quantum theory. &
\UIDT\ gives OSR a concrete physical carrier through the field $\Sx$. \\
\midrule
Informational structural realism & Reality is analysed through informational
objects and relational structures. & Connects ontology with information theory. &
\UIDT\ treats $\Sx$ as a dynamic information-density primitive, not as a digital
object or computational substrate. \\
\midrule
Strong emergentism & Higher-level phenomena have irreducible causal powers. &
Captures the apparent novelty of life, mind, and agency. &
\UIDT\ rejects strong emergence; complex phenomena are weakly emergent patterns
inside the causal dynamics of $\Sx$. \\
\midrule
Physicalism or materialism & The physical is fundamental. & Provides continuity
with the natural sciences. & \UIDT\ is an informational physicalism: matter is
derived, while the physically realised primitive is $\Sx$. \\
\bottomrule
\end{tabular}}
\caption{Philosophical positions in relation to UIDT.}
\label{tab:phil-comparison}
\end{table}
\newpage
%=============================================================================
\thispagestyle{uidtontology}
\vspace*{\fill}
\begin{center}
\begin{minipage}{0.72\textwidth}
\centering
{\color{darkgray!70}\rule{0.42\textwidth}{0.35pt}}\\[2.2em]
{\Large\itshape
As soon as there is distinction, there is being.\\[0.35em]
As soon as being is sufficiently complex, there is thought.\par}
\vspace{2.2em}
{\color{darkgray!45}\rule{0.18\textwidth}{0.3pt}}\\[1.6em]
{\footnotesize\scshape Original German}\\[0.7em]
{\normalsize\itshape
Sobald Unterscheidung ist, ist Sein.\\[0.25em]
Sobald Sein hinreichend komplex ist, ist Denken.\par}
\vspace{2.2em}
{\color{darkgray!70}\rule{0.42\textwidth}{0.35pt}}
\end{minipage}
\end{center}
\vspace*{\fill}
\newpage
%=============================================================================
% BIBLIOGRAPHY (Erweitert – integriert Research-Papiere und Desiderate)
%=============================================================================
\begin{thebibliography}{99}

\bibitem{Jaffe2000}
A.~Jaffe and E.~Witten,
``Quantum Yang--Mills Theory,''
Clay Mathematics Institute Millennium Problem Description, 2000.

% UIDT Core References
\bibitem{Rietz2026_Framework}
P.~Rietz,
``Vacuum Information Density as the Fundamental Geometric Scalar,''
UIDT Framework v3.9 (2026).
DOI: \href{https://doi.org/10.5281/zenodo.17835200}{10.5281/zenodo.17835200}.

\bibitem{Rietz2026_BigBang}
P.~Rietz,
``Big Bang Inflation: Relationale Informationsdichte als Phasenübergang der
ersten Unterscheidung,''
UIDT Research Report (2026).
[Internal note – to be submitted to arXiv].

\bibitem{Rietz2026_DeepScan}
P.~Rietz,
``UIDT Academic Deep Scan: Inflation Distinktion – Externe Validierung und
Lückenanalyse,''
UIDT Internal Audit Document (2026).
[Internal note – to be submitted to arXiv].

\bibitem{Rietz2026_Zusammenfassung}
P.~Rietz,
``Zusammenfassung: Prägeometrische Ansätze und die Einordnung von UIDT,''
UIDT Comparative Survey (2026).
[Internal note – to be submitted to arXiv].

% Foundational Texts
\bibitem{SpencerBrown1969}
G.~Spencer-Brown,
\emph{Laws of Form} (Allen \& Unwin, London, 1969).

\bibitem{Wheeler1990}
J.~A.~Wheeler,
``Information, Physics, Quantum: The Search for Links,''
in \emph{Complexity, Entropy, and the Physics of Information},
ed.\ W.~H.~Zurek (Addison-Wesley, 1990).

\bibitem{Lawvere1964}
F.~W.~Lawvere,
``An elementary theory of the category of sets,''
\emph{Proc.\ Natl.\ Acad.\ Sci.\ USA} \textbf{52}, 1506 (1964).

\bibitem{Leibniz1714}
G.~W.~Leibniz,
\emph{Monadologie} (1714).

\bibitem{Tegmark2008}
M.~Tegmark,
``The Mathematical Universe,''
\emph{Found.\ Phys.} \textbf{38}, 101 (2008).
arXiv: \href{https://arxiv.org/abs/0704.0646}{0704.0646}.

\bibitem{Shapiro1997}
S.~Shapiro,
\emph{Philosophy of Mathematics: Structure and Ontology}
(Oxford University Press, Oxford, 1997).

\bibitem{LeNepvou2025StructuralActualism}
A.~Le~Nepvou,
``Structural Realism Beyond Static Form:
Toward a Dynamically Grounded Ontology,''
Zenodo preprint, Version~v2 (2025).
doi: \href{https://doi.org/10.5281/zenodo.16962526}{10.5281/zenodo.16962526}.

% Pregeometric Literature
\bibitem{Sorkin2005}
R.~D.~Sorkin,
``Causal Sets: Discrete Gravity,''
in \emph{Lectures on Quantum Gravity}, eds.\ A.~Gomberoff and D.~Marolf
(Springer, 2005), pp.~305--327.
arXiv: \href{https://arxiv.org/abs/gr-qc/0309009}{gr-qc/0309009}.

\bibitem{Rovelli2004}
C.~Rovelli,
\emph{Quantum Gravity} (Cambridge University Press, Cambridge, 2004).

\bibitem{Swingle2012}
B.~Swingle,
``Entanglement Renormalization and Holography,''
\emph{Phys.\ Rev.\ D} \textbf{86}, 065007 (2012).
arXiv: \href{https://arxiv.org/abs/0905.1317}{0905.1317}.

\bibitem{Maldacena2013}
J.~Maldacena and L.~Susskind,
``Cool horizons for entangled black holes,''
\emph{Fortschr.\ Phys.} \textbf{61}, 781 (2013).
arXiv: \href{https://arxiv.org/abs/1306.0533}{1306.0533}.

\bibitem{Konopka2008}
T.~Konopka, F.~Markopoulou, and S.~Severini,
``Quantum Graphity: a model of emergent locality,''
\emph{Phys.\ Rev.\ D} \textbf{77}, 104029 (2008).
arXiv: \href{https://arxiv.org/abs/0801.0861}{0801.0861}.

% Philosophy of Science and Metaphysics
\bibitem{Heidegger1929}
M.~Heidegger,
\emph{Was ist Metaphysik?} (1929).
In \emph{Wegmarken} (Vittorio Klostermann, Frankfurt a.M., 1976).

\bibitem{Bateson1972}
G.~Bateson,
\emph{Steps to an Ecology of Mind} (Ballantine Books, New York, 1972).

\bibitem{Landauer1961}
R.~Landauer,
``Irreversibility and Heat Generation in the Computing Process,''
\emph{IBM J.\ Res.\ Dev.} \textbf{5}, 183 (1961).

\bibitem{Spinoza1677}
B.~Spinoza,
\emph{Ethica, ordine geometrico demonstrata} (1677).

\bibitem{Wigner1960}
E.~P.~Wigner,
``The Unreasonable Effectiveness of Mathematics in the Natural Sciences,''
\emph{Commun.\ Pure Appl.\ Math.} \textbf{13}, 1 (1960).

\bibitem{Bostrom2003}
N.~Bostrom,
``Are You Living in a Computer Simulation?''
\emph{Philos.\ Q.} \textbf{53}, 243 (2003).

\bibitem{Goedel1931}
K.~G\"odel,
``\"Uber formal unentscheidbare S\"atze der Principia Mathematica
und verwandter Systeme~I,''
\emph{Monatsh.\ Math.\ Phys.} \textbf{38}, 173 (1931).

\bibitem{Kuhn1962}
T.~S.~Kuhn,
\emph{The Structure of Scientific Revolutions}
(University of Chicago Press, Chicago, 1962).

\bibitem{Popper1934}
K.~R.~Popper,
\emph{Logik der Forschung} (Julius Springer, Vienna, 1934);
English edition: \emph{The Logic of Scientific Discovery}
(Hutchinson, London, 1959).

\bibitem{Popper1963}
K.~R.~Popper,
\emph{Conjectures and Refutations: The Growth of Scientific Knowledge}
(Routledge, London, 1963).

% Physics and Cosmology
\bibitem{Penrose2004}
R.~Penrose,
\emph{The Road to Reality} (Jonathan Cape, London, 2004).

\bibitem{Hawking1988}
S.~W.~Hawking,
\emph{A Brief History of Time} (Bantam Books, 1988).

\bibitem{DeWitt1967}
B.~S.~DeWitt,
``Quantum Theory of Gravity. I. The Canonical Theory,''
\emph{Phys.\ Rev.} \textbf{160}, 1113 (1967).

% Lattice QCD and Yang-Mills
\bibitem{Chen2006}
Y.~Chen \emph{et al.},
``Glueball spectrum and matrix elements on anisotropic lattices,''
\emph{Phys.\ Rev.\ D} \textbf{73}, 014516 (2006).
arXiv: \href{https://arxiv.org/abs/hep-lat/0510074}{hep-lat/0510074}.

\bibitem{DESI2025}
DESI Collaboration,
``DESI DR2 Baryon Acoustic Oscillation Measurements,''
arXiv: \href{https://arxiv.org/abs/2503.14738}{2503.14738} (2025).

\bibitem{Planck2018}
Planck Collaboration,
``Planck 2018 results. VI. Cosmological parameters,''
\emph{Astron.\ Astrophys.} \textbf{641}, A6 (2020).
arXiv: \href{https://arxiv.org/abs/1807.06209}{1807.06209}.

\bibitem{Weinberg1995}
S.~Weinberg,
\emph{The Quantum Theory of Fields, Vol.~I: Foundations}
(Cambridge University Press, Cambridge, 1995).

\bibitem{Shifman1979}
M.~A.~Shifman, A.~I.~Vainshtein, and V.~I.~Zakharov,
``QCD and Resonance Physics,''
\emph{Nucl.\ Phys.\ B} \textbf{147}, 385 (1979).

\bibitem{Huang1999}
T.~Huang \emph{et al.},
``Scalar glueball in QCD sum rules with instanton corrections,''
\emph{Phys.\ Rev.\ D} \textbf{59}, 034026 (1999).

\bibitem{Banach1922}
S.~Banach,
``Sur les op\'erations dans les ensembles abstraits et leur application
aux \'equations int\'egrales,''
\emph{Fund.\ Math.} \textbf{3}, 133 (1922).

\bibitem{Jacobson1995}
T.~Jacobson,
``Thermodynamics of Spacetime: The Einstein Equation of State,''
\emph{Phys.\ Rev.\ Lett.} \textbf{75}, 1260 (1995).
arXiv: \href{https://arxiv.org/abs/gr-qc/9504004}{gr-qc/9504004}.

\bibitem{VanRaamsdonk2010}
M.~Van Raamsdonk,
``Building up spacetime with quantum entanglement,''
\emph{Gen.\ Rel.\ Grav.} \textbf{42}, 2323 (2010).
arXiv: \href{https://arxiv.org/abs/1005.3035}{1005.3035}.

\bibitem{Verlinde2011}
E.~Verlinde,
``On the Origin of Gravity and the Laws of Newton,''
\emph{JHEP} \textbf{1104}, 029 (2011).
arXiv: \href{https://arxiv.org/abs/1001.0785}{1001.0785}.

\bibitem{Chishtie2025EmergentSpacetime}
F.~A.~Chishtie,
``Emergent spacetime from spatial energy potentiality:
a new theoretical framework for early universe cosmology,''
arXiv: \href{https://arxiv.org/abs/2502.18524}{2502.18524}
[physics.gen-ph] (2025; rev. 2026).

\bibitem{SumayaMartinez2026FisherTime}
J.~Sumaya-Martinez,
``Fisher-Informational Time:
A Causal-Geometric Framework for Emergent Clock Time Physical Distinguishability,''
arXiv: \href{https://arxiv.org/abs/2605.03958}{2605.03958}
[quant-ph] (2026).

\bibitem{Zeilinger1999}
A.~Zeilinger,
``A Foundational Principle for Quantum Mechanics,''
\emph{Found.\ Phys.} \textbf{29}, 631 (1999).

\bibitem{LadymanRoss2007}
J.~Ladyman and D.~Ross,
\emph{Every Thing Must Go: Metaphysics Naturalized}
(Oxford University Press, Oxford, 2007).

\bibitem{Floridi2008}
L.~Floridi,
``A defence of informational structural realism,''
\emph{Synthese} \textbf{161}, 219--253 (2008).
doi: \href{https://doi.org/10.1007/s11229-007-9163-z}{10.1007/s11229-007-9163-z}.

\bibitem{Gudder2015}
S.~Gudder,
``Discrete Quantum Gravity,''
arXiv: \href{https://arxiv.org/abs/1507.01281}{1507.01281} (2015).

\bibitem{Bekenstein1973}
J.~D.~Bekenstein,
``Black Holes and Entropy,''
\emph{Phys.\ Rev.\ D} \textbf{7}, 2333 (1973).

\bibitem{Padmanabhan2010}
T.~Padmanabhan,
``Thermodynamical Aspects of Gravity: New Insights,''
\emph{Rep.\ Prog.\ Phys.} \textbf{73}, 046901 (2010).
arXiv: \href{https://arxiv.org/abs/0911.5004}{0911.5004}.

\bibitem{Athenodorou2021}
A.~Athenodorou and M.~Teper,
``The glueball spectrum of SU(3) gauge theory in 3+1 dimensions,''
\emph{JHEP} \textbf{2021}, 172 (2021).
arXiv: \href{https://arxiv.org/abs/2007.06422}{2007.06422}.

\bibitem{Morningstar1999}
C.~Morningstar and M.~Peardon,
``The glueball spectrum from an anisotropic lattice study,''
\emph{Phys.\ Rev.\ D} \textbf{60}, 034509 (1999).
arXiv: \href{https://arxiv.org/abs/hep-lat/9901004}{hep-lat/9901004}.

\bibitem{RyuTakayanagi2006}
S.~Ryu and T.~Takayanagi,
``Holographic Derivation of Entanglement Entropy from the anti-de Sitter
Space/Conformal Field Theory Correspondence,''
\emph{Phys.\ Rev.\ Lett.} \textbf{96}, 181602 (2006).
arXiv: \href{https://arxiv.org/abs/hep-th/0603001}{hep-th/0603001}.

\bibitem{Vadacchino2024}
D.~Vadacchino, A.~Nada, and C.~Bonanno,
``Topological susceptibility of SU(3) pure-gauge theory from out-of-equilibrium simulations,''
\emph{PoS} \textbf{LATTICE2024}, 415 (2024).
doi: \href{https://doi.org/10.22323/1.466.0415}{10.22323/1.466.0415};
arXiv: \href{https://arxiv.org/abs/2411.00620}{2411.00620}.

\end{thebibliography}

%=============================================================================

%=============================================================================

%=============================================================================
\end{document}
