% ============================================================
% UIDT v3.9 --- Clay Submission Appendix
% Monte Carlo Validation Evidence
% ============================================================
% Author: P. Rietz
% ORCID: 0009-0007-4307-1609
% Canonical DOI: 10.5281/zenodo.17835200
% Dataset DOI:   10.5281/zenodo.17554179
% Date:          2026-04-29
%
% EVIDENCE SYSTEM:
%   [A]  mathematically proven
%   [A-] phenomenological parameter
%   [B]  lattice compatible
%   [C]  calibrated cosmology
%   [D]  prediction
%   [E]  speculative
%
% LANGUAGE RULE: no prestige/closure wording.
% OPEN ITEMS are retained verbatim and must not be removed.
% ============================================================

\section*{Appendix~MC: Monte Carlo Validation Evidence}
\label{app:mc_evidence}

This appendix documents the statistical validation of the UIDT~v3.9 canonical
parameter set against the Monte Carlo chain of $N=100\,000$ thinned samples
(DOI~\href{https://doi.org/10.5281/zenodo.17554179}{10.5281/zenodo.17554179}).
All evidence categories follow the UIDT Evidence System (A--E).

%------------------------------------------------------------
\subsection*{MC.1 MCMC Convergence Diagnostics \textnormal{\textmd{[A]}}}
%------------------------------------------------------------

All ten sampled parameters satisfy publication-level convergence criteria
(Gelman--Rubin split-$\hat{R}<1.001$; effective sample size $N_{\rm eff}>96\,000$;
integrated autocorrelation time $\tau_{\rm int}\approx 0.50$, indicating
near-fully decorrelated samples after thinning).

\begin{table}[h]
\centering
\caption{MCMC convergence summary (all 10 parameters).}
\begin{tabular}{lcccc}
\hline
Parameter & Split-$\hat{R}$ & $N_{\rm eff}$ & $\tau_{\rm int}$ & Verdict \\
\hline
$\Delta^*$ & $\approx1.000$ & $>96\,000$ & $\approx0.50$ & converged \\
$\gamma$   & $\approx1.000$ & $>96\,000$ & $\approx0.50$ & converged \\
$\Psi$     & $\approx1.000$ & $>96\,000$ & $\approx0.50$ & converged \\
$\kappa$   & $\approx1.000$ & $>96\,000$ & $\approx0.50$ & converged \\
$\lambda_S$& $\approx1.000$ & $>96\,000$ & $\approx0.50$ & converged \\
$m_S$      & $\approx1.000$ & $>96\,000$ & $\approx0.50$ & converged \\
\text{kinetic\_VEV} & $\approx1.000$ & $>96\,000$ & $\approx0.50$ & converged \\
$\Pi_S$    & $\approx1.000$ & $>96\,000$ & $\approx0.50$ & converged \\
$w_0$      & $\approx1.000$ & $>96\,000$ & $\approx0.50$ & converged \\
$E_T$      & $\approx1.000$ & $>96\,000$ & $\approx0.50$ & converged \\
\hline
\end{tabular}
\end{table}

%------------------------------------------------------------
\subsection*{MC.2 LEDGER Consistency \textnormal{\textmd{[A] / [A-]}}}
%------------------------------------------------------------

Table~\ref{tab:ledger_consistency} summarises the consistency of the MC
posterior means with the UIDT immutable parameter ledger.

\begin{table}[h]
\centering
\caption{MC posterior vs.\ LEDGER.}
\label{tab:ledger_consistency}
\begin{tabular}{lcccccc}
\hline
Param. & LEDGER & MC mean & MC $\sigma$ & $|$residual$|$ & $z$ & $p$ \\
\hline
$\Delta^*$ [GeV] & $1.710\pm0.015$ & $1.710044$ & $0.014993$
  & $4.44\times10^{-5}$ & $0.003$ & $0.9976$ \\
$\gamma$ & $16.339$ & $16.37395$ & $1.00513$
  & $0.035$ & $-0.0348$ & $0.9723$ \\
\hline
\end{tabular}
\end{table}

Both parameters are consistent with their LEDGER values at the $<0.04\,\sigma$
level.
For $\Delta^*$ the evidence category is [A];
for $\gamma$ it is [A-] because the LEDGER reference is phenomenologically
determined, not first-principles derived.

%------------------------------------------------------------
\subsection*{MC.3 Holographic Dressing Relation
  \textnormal{\textmd{[B/D] $\to$ [C]}}}
%------------------------------------------------------------

The bare-gamma extrapolation (\verb|docs/bare_gamma_theorem.md|) gives
$\gamma_\infty = 16.3437\pm10^{-4}$~[B] with dressing shift
$\delta\gamma = 0.0047$~[B/D].
Amplified by the holographic volume factor $L^4$ ($L=8.2$, non-canonical
free parameter, open issue~S1-01):
\[
  w_a = -\frac{\delta\gamma}{\gamma_\infty}\,L^4
        \approx -1.300 \quad [C]
\]
The DESI~DR2 + Union3 reference value is
$w_a = -1.27\pm0.37$ (arXiv:2503.14738);
the Mahalanobis distance is $0.65\,\sigma$, indicating compatibility.

\textbf{Limitation:} $L=8.2$ is not a canonical UIDT parameter.
All $w_a$ values derived via the $L^4$ amplification inherit this ambiguity
(issue~S1-01 unresolved).
H$_0$ and $S_8$ tensions are \textbf{not} declared resolved.

%------------------------------------------------------------
\subsection*{MC.4 Lattice QCD Compatibility \textnormal{\textmd{[B]}}}
%------------------------------------------------------------

The quenched-approximation scalar glueball mass from
Morningstar \& Peardon~\cite{Morningstar1999}:
$\Delta^*(\text{lattice}) = 1.730\pm0.090$~GeV.

Comparison with the MC result $\Delta^*(\text{MC}) = 1.710044\pm0.014993$~GeV:
\[
  \chi^2 = 0.0478, \quad p = 0.8269, \quad |z| = 0.2187\,\sigma
\]
Full compatibility is indicated.
Evidence category: [B] (numerical extrapolation, quenched approximation;
not an independent first-principles confirmation).

\begin{thebibliography}{9}
\bibitem{Morningstar1999}
  C.~Morningstar and M.~Peardon,
  Phys.\ Rev.\ D~\textbf{60}, 034509 (1999).
  DOI~\href{https://doi.org/10.1103/PhysRevD.60.034509}{10.1103/PhysRevD.60.034509};
  arXiv:hep-lat/9901004.
\end{thebibliography}

%------------------------------------------------------------
\subsection*{MC.5 Anti-Correlation $r(\gamma, \text{kinetic\_VEV})$
  \textnormal{\textmd{[B]}}}
%------------------------------------------------------------

The stored correlation matrix reports
$r(\gamma, \text{kinetic\_VEV}) = -0.982130546$.
This is consistent with an anti-proportional coupling and with the $2.3$-factor
damping mapping documented in Appendix~III of the main manuscript.
Evidence category: [B] (stored artefact, not recomputed from raw chain in this version).

%------------------------------------------------------------
\subsection*{MC.6 Open Item~1: Provenance Conflict
  $r(\Delta^*, \Pi_S)$ \textnormal{\textmd{\color{red}[TENSION ALERT]}}}
%------------------------------------------------------------

\textbf{[TENSION ALERT]}

The stored correlation matrix reports
$r(\Delta^*, \Pi_S) = +0.720284420$.
The raw-chain audit session explicitly states that this value was
\emph{not reproduced} from \verb|UIDT_MonteCarlo_samples_100k.csv|.

\begin{center}
\begin{tabular}{lcc}
\hline
Source & $r(\Delta^*, \Pi_S)$ \\
\hline
Stored correlation matrix & $+0.720284420$ \\
Raw-chain audit & not reproduced \\
\hline
\end{tabular}
\end{center}

This conflict must be resolved by a dedicated script that recomputes all
45~pairwise correlations directly from the raw CSV and compares them
against the stored matrix before any evidence-category upgrade
($\mathrm{A}^-\!\to\mathrm{A}$) is considered.

%------------------------------------------------------------
\subsection*{MC.7 Open Item~2: Soft RG Residual
  \textnormal{\textmd{\color{red}[RG\_CONSTRAINT\_FAIL]}}}
%------------------------------------------------------------

\textbf{[RG\_CONSTRAINT\_FAIL]}

The RG fixed-point relation $5\kappa^2 = 3\lambda_S$ is satisfied at the
posterior-mean level (evidenced by $r(\kappa, \lambda_S)\approx+0.99$),
but the residual $|5\kappa^2 - 3\lambda_S|$ computed from the
high-precision mean values in
\verb|UIDT_HighPrecision_mean_values.csv| was flagged as exceeding
the formal tolerance $1\times10^{-14}$ set in
\verb|verification/tests/test_monte_carlo_summary.py|.

This flag must \textbf{not} be silently removed or smoothed over.
A numerical correction to the hp-mean values, or a documented revision
of the tolerance, is required before this item is closed.

% ============================================================
% END OF APPENDIX MC
% ============================================================
